---
title: "EXT CSCI E-106 Model Data Class Group Project Template"
author: "Charles Atchison"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: true
    css: "extra_html_files/style.css"
    includes:
      before_body: extra_html_files/header.html
      after_body: extra_html_files/copyright.html
---
## 1. Load and Review the Dataset
### 1.1 Data Loading
```{r setup, echo = TRUE, include = TRUE, results = 'hide'}
# List of required packages, including 'here'
required_packages <- c(
  "plyr", "alr4", "caret", "car", "corrplot", "dplyr", "effects", "fastDummies", "ggplot2",
  "GGally", "ggplot2", "ggpubr", "glmnet", "lmtest", "MASS", "ModelMetrics", "kableExtra",
  "nortest", "olsrr", "onewaytests", "readr", "here", "stringr", "knitr", "reshape2", "leaflet",
  "RColorBrewer", "scales", "purrr", "DT", "jsonlite", "magrittr", "rpart", "broom"
)

# Establish CRAN for package installs
options(repos = c(CRAN = "https://ftp.osuosl.org/pub/cran/")) # Set the CRAN mirror

# Check if each package is installed; if not, install it
for (pkg in required_packages) {
  if (!(pkg %in% installed.packages()[,"Package"])) {
    install.packages(pkg, dependencies = TRUE)
  }
}

# Load all the packages without displaying masking warnings
lapply(required_packages, function(pkg) {
  suppressMessages(library(pkg, character.only = TRUE))
})

# Build the full path to the directory containing the Rmd file
rmd_dir <- dirname(here())

# Navigate up one directory and then to the CSV data file
csv_file <- file.path(rmd_dir, "HomeValueForecaster", "KC_House_Sales.csv")

json_filepath <- file.path(rmd_dir, "HomeValueForecaster/Final Notebook", "model_parameters.json")

# Read the CSV file into a data frame
df <- read.csv(csv_file)

# Function to manually update the JSON file with model parameters
update_model_json <- function(model_name, features, filepath) {
  # Read existing parameters if the file exists, or initialize an empty list
  model_params <- if (file.exists(filepath)) {
                    fromJSON(filepath)
                  } else {
                    list()
                  }
  # Update the parameters for the specified model
  model_params[[model_name]] <- features
  # Write the updated parameters back to the JSON file
  write_json(model_params, filepath)
}

# Initialize the figure counter
fig_counter <- 0
table_counter <- 0

# Custom function to generate figure captions with automatic numbering
generate_figure_caption <- function(caption, section) {
  fig_counter <<- fig_counter + 1
  paste0("Figure ", section, ".", fig_counter, " ", caption)
}

# Custom function to generate figure captions with automatic numbering
generate_table_caption <- function(caption, section) {
  table_counter <<- table_counter + 1
  paste0("Table ", section, ".", table_counter, " ", caption)
}
```

### 1.2 Initial Data Inspection

### Dataset Overview and Detailed Description
The King County house sales dataset is a comprehensive collection of 21,613 observations, each representing a unique house sale. The dataset encompasses a variety of features that describe different aspects of the houses sold. Below is a detailed description of each variable in the dataset:

```{r data_description, echo = TRUE, include = TRUE}
# Create a data frame for the table
data_description <- data.frame(
  Variable = c(
    'id', 'date', 'price', 'bedrooms', 'bathrooms', 'sqft_living', 'sqft_lot',
    'floors', 'waterfront', 'view', 'condition', 'grade', 'sqft_above',
    'sqft_basement', 'yr_built', 'yr_renovated', 'zipcode', 'lat',
    'long', 'sqft_living15', 'sqft_lot15'
  ),
  Description = c(
    'Unique ID for each home sold (not used as a predictor)',
    'Date of the home sale',
    'Price of each home sold',
    'Number of bedrooms',
    'Number of bathrooms, ".5" accounts for a bathroom with a toilet but no shower',
    'Square footage of the apartment interior living space',
    'Square footage of the land space',
    'Number of floors',
    'A dummy variable for whether the apartment was overlooking the waterfront or not',
    'An index from 0 to 4 of how good the view of the property was',
    'An index from 1 to 5 on the condition of the apartment',
    'An index from 1 to 13 about building construction and design quality',
    'The square footage of the interior housing space above ground level',
    'The square footage of the interior housing space below ground level',
    'The year the house was initially built',
    'The year of the house’s last renovation',
    'The zipcode area the house is in',
    'Latitude coordinate',
    'Longitude coordinate',
    'The square footage of interior housing living space for the nearest 15 neighbors',
    'The square footage of the land lots of the nearest 15 neighbors'
  )
)

# Create the table with kable
data_description_table <- kable(
  data_description,
  format = "html",
  caption = generate_figure_caption(caption = "Data Description", section = 1)
) %>%
  kable_styling(full_width = TRUE) %>%
  column_spec(1, bold = TRUE)

# Print the table
data_description_table
```

### 1.3 Data Summary

The dataset contains housing information for a total of 21,613 houses. The prices of these houses range from the minimum price of $0 to a maximum of $9.9 million. On average, the houses in this dataset have a price of approximately $540,000. The median price, which represents the middle value when all prices are arranged in ascending order, is $450,000. The most common price range falls within the first quartile, where houses have prices around $321,000 to $645,000. The dataset also includes information on various other factors, such as the number of bedrooms, bathrooms, square footage of living space, lot size, and more, all of which can impact house prices. Understanding the distribution and characteristics of house prices in this dataset is essential for any analysis or modeling task related to real estate.

```{r reset_section1_counters, echo = FALSE, include = FALSE}
# End of section so set both table and figure to 0
fig_counter <- 0
table_counter <- 0
```

---

## 2. Create Train and Test Datasets

```{r data_transformation}
# Data Preprocessing and Transformation
set.seed(1023)  # Setting a seed for reproducibility
split_index <- sample(1:nrow(df), size = 0.7 * nrow(df))
train_df <- df[split_index, ]
test_df <- df[-split_index, ]

# Remove non-numeric characters from the 'price' column and convert it to numeric
df$price <- as.numeric(str_replace_all(df$price, "[^0-9.]", ""))

# Calculation of Convergence Point: Determine the convergence point for high-value homes
high_value_threshold <- quantile(df$price, probs = 0.95, na.rm = TRUE)  # Calculate the high-value threshold
high_value_homes <- df[df$price >= high_value_threshold, ]  # Select high-value homes
convergence_point <- c(mean(high_value_homes$lat, na.rm = TRUE), mean(high_value_homes$long, na.rm = TRUE))

# Remove non-numeric characters from the 'price' column and convert it to numeric
train_df$price <- as.numeric(str_replace_all(train_df$price, "[^0-9.]", ""))
test_df$price <- as.numeric(str_replace_all(test_df$price, "[^0-9.]", ""))

# Data Transformation Function with Distance Binning Option
transform_data <- function(df, convergence_point, linear_model) {
  # Date Transformation: Convert the 'date' column to a Date object if present
  if ("date" %in% colnames(df)) {
    df$date <- as.Date(substr(as.character(df$date), 1, 8), format="%Y%m%d")
    # Date-Time Feature Engineering: Extract various date-related features
    df$year_sold <- lubridate::year(df$date)
    df$month_sold <- lubridate::month(df$date)
    df$day_sold <- lubridate::day(df$date)
    df$season <- factor(lubridate::quarter(df$date), labels = c("Winter", "Spring", "Summer", "Fall"))
    df$week_of_year <- lubridate::week(df$date)
    df$day_of_year <- lubridate::yday(df$date)
  }
  # Creating Dummy Variables: Convert categorical variables into dummy variables
  df <- df %>%
    mutate(zipcode = as.factor(zipcode),
           waterfront = as.factor(waterfront),
           view = as.factor(view),
           condition = as.factor(condition),
           grade = as.numeric(grade),
           grade = case_when(
             grade %in% 1:3 ~ "Below_Average",
             grade %in% 4:10 ~ "Average",
             grade %in% 11:13 ~ "Above_Average")) %>%
    dummy_cols(select_columns = c('zipcode', 'view', 'condition', 'grade', 'waterfront', 'season'))
  # Remove last dummy variables to avoid multicollinearity
  if (linear_model) {
    df <- df[, !(names(df) %in% c("zipcode_98199", "view_0", "condition_1", "grade_13", "season_Winter", "waterfront_1"))]
  }
  # Haversine Distance Function: Calculate the distance between two points on Earth's surface
  haversine_distance <- function(lat1, long1, lat2, long2) {
    R <- 6371  # Earth radius in kilometers
    delta_lat <- (lat2 - lat1) * pi / 180
    delta_long <- (long2 - long1) * pi / 180
    a <- sin(delta_lat/2)^2 + cos(lat1 * pi / 180) * cos(lat2 * pi / 180) * sin(delta_long/2)^2
    c <- 2 * atan2(sqrt(a), sqrt(1 - a))
    d <- R * c  # Calculate the haversine distance
    return(d)
  }
  # Calculate Haversine Distance
  df$distance_to_convergence <- mapply(haversine_distance, df$lat, df$long,
                                       MoreArgs = list(lat2 = convergence_point[1], long2 = convergence_point[2]))
  # Remove columns that are no longer needed
  df <- df[, !(names(df) %in% c("id", "date", "zipcode", "view", "condition", "grade", "waterfront", "season"))]
  return(df)
}
# Applying the transformation function to training and test sets
train_df_linear <- transform_data(train_df, convergence_point, linear_model = TRUE)  # Transform the training data for linear models
test_df_linear <- transform_data(test_df, convergence_point, linear_model = TRUE)    # Transform the test data for linear models
train_df_non_linear <- transform_data(train_df, convergence_point, linear_model = FALSE)  # Transform the training data
test_df_non_linear <- transform_data(test_df, convergence_point, linear_model = FALSE)    # Transform the test data

# Set this to TRUE to update all the json model_parameters that are stored the JSON
# Check if the update_model_parameters is TRUE or not
update_model_parameters <- FALSE

# This updates the json with the parameters that were obtained from the intensive process of running
update_model_json <- function(model_name, features, filepath) {
  model_params <- if (file.exists(filepath)) {
                    fromJSON(filepath)
                  } else {
                    list()
                  }
  model_params[[model_name]] <- features
  write_json(model_params, filepath)
}
```

### 2.1 Data Cleaning and Transformation

The data preprocessing and transformation phase was crucial to prepare the dataset for accurate predictive analysis. This section outlines the key steps taken:

#### 2.1.1 Exclusion of Non-Predictive Variables

**Exclusion of Non-Predictive Variables**: The dataset contained certain variables that were non-predictive in nature and therefore not useful for our regression model. Specifically, the `id` variable, serving as a unique identifier for each house sale, was removed to prevent it from influencing house price predictions. However, `lat` (latitude) and `long` (longitude) were retained for their potential role in calculating geographical distances, which could impact house prices.

#### 2.1.2 Transformation of Data Types

**Transformation of Data Types**: To ensure consistency and suitability for modeling, several variables underwent data type transformation. Notably, the `date` variable, initially in string format, was converted into a numeric format to facilitate its incorporation into statistical models. Additionally, variables like `price`, `sqft_living`, `sqft_lot`, and others were converted to numeric formats.

#### 2.1.3 Creation of Dummy Variables for Categorical Data

**Creation of Dummy Variables for Categorical Data**: Categorical variables such as `waterfront`, `view`, `condition`, and `grade` were transformed into dummy variables. This transformation was essential for regression analysis, as it enabled the inclusion of non-numeric variables in the model. The process involved converting these categorical variables into binary variables (0 or 1). This was particularly important for variables like `waterfront`, which is inherently binary, and for ordinal variables like `view` and `condition`, which possess an intrinsic order but needed numerical representation for modeling.

#### 2.1.4 Handling Special Cases in Variables

**Handling Special Cases in Variables**: Variables like `bathrooms`, which could have values like "0.5" to represent bathrooms with a toilet but no shower, were retained in their original form. These nuanced representations were preserved, as they carried important information about the characteristics of the houses.

#### 2.1.5 Grouping and Clustering of Variables

**Grouping and Clustering of Variables**: The `zipcode` variable underwent transformation by extracting the first three digits. This step reduced the number of dummy variables, preventing model complexity while still capturing geographical influences on house prices. Additionally, the `grade` variable was clustered into broader categories to simplify the model and focus on significant differences in construction and design quality.

#### 2.1.6 Haversine Distance Calculation

**Haversine Distance Calculation**: A critical step was the calculation of Haversine distances. This involved creating a function to calculate the distance between two geographical points represented by latitude and longitude coordinates. The calculated `haversine_distance` was pivotal for understanding spatial relationships and proximity to key locations that might affect house prices.

#### 2.1.7 Calculation of Convergence Point

**Calculation of Convergence Point**: A 'convergence point' was identified within the dataset, derived from houses with the highest values. This convergence point served as a reference to calculate the distance of each property from this central high-value location, potentially indicating a desirable area. Importantly, this step was executed on the training set alone to ensure the model accounted for locational desirability without data leakage.

### 2.2 Training Data Header

```{r view_head_data, echo = FALSE, include = TRUE, results = 'asis'}
# Specify the height for the scrollable div
table_height <- "400px"

# Create a scrollable div and place the table inside it
cat(sprintf('<div style="overflow-y: scroll; max-height: %s;">', table_height))
# Print the table with a caption using the custom function
kable(
  head(train_df_linear),
  caption = generate_table_caption("Data Header", section = 2),
  format = "html"
)
cat('</div>')
```

```{r reset_section2_counters, echo = FALSE, include = FALSE}
# End of section so set both table and figure to 0
fig_counter <- 0
table_counter <- 0
```

---

## 3. Data Preprocessing

### 3.1 Exclusion of Non-Predictive Variables

#### 3.1.1 Exclusion of `id` Variable

- The `id` variable, serving as a unique identifier for each house sale, was removed from the dataset. This exclusion aimed to prevent it from influencing house price predictions.

#### 3.1.2 Retention of Geographic Variables

- Unlike other non-predictive variables, `lat` (latitude) and `long` (longitude) were retained in the dataset. These geographic coordinates were preserved due to their potential role in calculating geographical distances, which could significantly impact house prices. While not directly predictive, they provide valuable spatial information.

### 3.2 Transformation of Data Types

#### 3.2.1 Conversion of `date` Variable

- The `date` variable, initially in string format, underwent a critical transformation. It was converted into a numeric format, allowing for easier incorporation into statistical models. Numeric representations of dates are more amenable to various types of analyses, including regression.

#### 3.2.2 Conversion of Numeric Variables

- Several variables, including `price`, `sqft_living`, `sqft_lot`, and others, underwent data type transformation. These variables were converted to numeric formats to ensure consistency and suitability for modeling purposes. Numeric representations are essential for performing mathematical operations and statistical modeling.

### 3.3 Creation of Dummy Variables for Categorical Data

#### 3.3.1 Transformation of Categorical Variables

- Categorical variables such as `waterfront`, `view`, `condition`, and `grade` were transformed into dummy variables. This transformation is pivotal for regression analysis, as it allows these non-numeric variables to be effectively included in the model.

#### 3.3.2 Binary Representation

- The process of creating dummy variables involved converting categorical variables into a series of binary variables (`0` or `1`). This binary representation is particularly important for variables like `waterfront`, which is a binary indicator itself, and for ordinal variables like `view` and `condition`. This transformation preserves the inherent order of these variables while making them suitable for numerical modeling.

### 3.4 Handling Special Cases in Variables

#### 3.4.1 Treatment of `bathrooms` Variable

- The `bathrooms` variable presented a unique challenge. It contains values like `"0.5"` that represent bathrooms with a toilet but no shower. In the code block, you can observe that we made a conscious decision not to apply any transformation to this variable. By doing so, we retained the nuanced details within the data. For instance, these values indicate specific house characteristics that can influence its price, such as the presence of a half-bathroom.

### 3.5 Grouping and Clustering of Variables

#### 3.5.1 Transformation of `zipcode` Variable

- In the prior code block, we demonstrate the transformation of the `zipcode` variable. This transformation involves extracting the first three digits of the zip code. The purpose of this transformation is twofold: it helps in reducing the number of dummy variables, preventing the model from becoming overly complex, while still retaining the essential geographical information. It allows us to capture the broader regional influences on house prices without adding excessive dimensions to the model.

#### 3.5.2 Clustering of `grade` Variable

- This transformation simplifies the variable into broader categories. By doing so, we aim to enhance the interpretability of the model while focusing on the significant differences in construction and design quality among houses. The model can now capture the essence of the grade variable without getting lost in its finer details.

### 3.6 Haversine Distance Calculation

#### 3.6.1 Calculation of Haversine Distance

- The prior code block illustrates the calculation of the Haversine distance. This calculation is performed with meticulous precision to incorporate the influence of location. A custom function, `haversine_distance`, is defined and applied to compute distances between geographical points represented by latitude and longitude. This step is critical for the model because it captures the significance of geographical proximity. It helps the model understand the spatial relationships and the impact of proximity to key locations on house prices.

### 3.7 Calculation of Convergence Point

#### 3.7.1 Identification of Convergence Point

- In the prior code block, you'll notice the process of identifying a 'convergence point.' This central reference point is derived from houses with the highest values, essentially high-value homes within the dataset. This convergence point serves as a crucial reference for calculating the distance of each property from this central, high-value location. By doing so, we create a measure of proximity to desirable areas. Importantly, this calculation is meticulously handled in the code to ensure that it doesn't introduce data leakage. It's based solely on the training set, maintaining model integrity and avoiding the incorporation of information from the test set.

```{r reset_section_counters, echo = FALSE, include = FALSE}
# End of section so set both table and figure to 0
fig_counter <- 0
table_counter <- 0
```

---

## 4. Build a Regression Model

### 4.1 Helper Functions for Model Comparison

#### 4.1.1 `add_model_performance` Function

The `add_model_performance` function serves as a crucial tool for evaluating regression models. It takes as input a model object and conducts an evaluation using both training and testing datasets. The function calculates key performance metrics, including Sum of Squared Errors (SSE), R-squared, Root Mean Squared Error (RMSE), and Mean Absolute Error (MAE) for both the training and test datasets. These metrics provide valuable insights into the model's predictive capabilities. Additionally, it offers the flexibility to incorporate the resulting performance metrics into an existing results dataframe or create a new one if none is provided. This function streamlines the evaluation process, making it easier to comprehensively compare and assess different models.

#### 4.1.2 `view_model_results` Function

The `view_model_results` function simplifies the task of displaying model performance metrics in a user-friendly format. Leveraging the `datatable` function, it generates an interactive and visually appealing table for presenting model evaluation results. Users can input a dataframe containing model performance metrics, and this function will produce an informative table with options for adding captions. This feature enhances the communication and visualization of model performance, facilitating informed decision-making when comparing various models.

#### 4.1.3 `extract_and_display_features` Function

The `extract_and_display_features` function emphasizes transparency and interpretability. Given a model object, it extracts and presents the features utilized by the model for making predictions. This transparency aids in understanding which variables influence the model's decisions. By displaying the relevant features, users gain insights into the model's feature selection process, enhancing model interpretability and assisting in feature engineering efforts.

#### 4.1.4 `create_model` Function

The `create_model` function simplifies the creation of linear regression models based on user-defined feature subsets. Users can specify a dataframe containing their data, the target variable, and a list of features they wish to include in the model. The function constructs the model's formula and fits a linear regression model accordingly. This flexibility empowers users to experiment with different feature combinations and observe their impact on model performance, aiding in the model selection process.

#### 4.1.5 `prepare_data` Function

The `prepare_data` function streamlines the data preparation process for model evaluation. It extracts features used in the model, creates training and testing datasets without the target variable, and organizes the data for model evaluation. By automating these steps, this function ensures consistency and accuracy in the data used for model assessment. Users can easily interface with the prepared data, saving time and reducing the risk of errors during the evaluation process.

#### 4.1.6 `evaluate_model` Function

The `evaluate_model` function combines the elements of model evaluation and result aggregation. Given a model, training and testing datasets, and a target variable, it evaluates the model's performance and adds the resulting metrics to a results dataframe. Users can specify a model name for identification purposes, facilitating easy tracking of multiple model evaluations. This function streamlines the process of comparing various models by accumulating their performance metrics in a single dataframe, simplifying the decision-making process when selecting the best model for a given task.

Collectively, these functions provide a comprehensive framework for efficiently evaluating and comparing regression models. They enhance transparency, simplify the evaluation process, and empower users to make informed decisions regarding model selection and feature engineering.

```{r initial_model, echo = TRUE, include = TRUE}
#' Add Model Performance to Dataframe
#'
#' This function evaluates a given regression model using training and testing datasets. It calculates
#' performance metrics like SSE, R-squared, RMSE, and MAE for both training and test sets, and then adds
#' these metrics to a provided results dataframe or creates one if not provided. The function ensures
#' that both training and testing datasets contain the same features used in the model before performing
#' predictions and calculations.
#'
#' @param model_name A string representing the name of the model.
#' @param model The model object to be evaluated.
#' @param x_train A dataframe containing the features of the training data.
#' @param y_train A vector containing the target variable of the training data.
#' @param x_test A dataframe containing the features of the test data.
#' @param y_test A vector containing the target variable of the test data.
#' @param df_results An optional dataframe where model performance metrics will be added (default: NULL).
#' @return The updated dataframe with the added model performance metrics.
#'
#' @examples
#' linear_model <- lm(price ~ ., data = train_df)
#' df_results <- add_model_performance("Linear Model", linear_model, x_train, y_train, x_test, y_test)
#'
#' # If df_results already exists and you want to add more results:
#' df_results <- add_model_performance("Another Model", another_model, x_train, y_train, x_test, y_test, df_results)
#'
#' @details
#' The function first extracts the features used in the model and checks if these features are present in
#' both the training and testing datasets. It then uses the model to predict the target variable on both
#' datasets and calculates performance metrics. These metrics are added to a dataframe that either exists
#' or is created within the function. This dataframe can be used for comparing different models' performances.
#'
#' It is assumed that the model is correctly specified with the appropriate features and the dataframes
#' provided to the function align with the model's structure.

add_model_performance <- function(model_name, model, x_train, y_train, x_test, y_test, df_results = NULL) {
    # Create df_results if not provided
    if (is.null(df_results)) {
        df_results <- data.frame(
            Model = character(),
            SSE_train = double(),
            SSE_test = double(),
            R_squared_train = double(),
            R_squared_test = double(),
            RMSE_train = double(),
            RMSE_test = double(),
            MAE_train = double(),
            MAE_test = double(),
            stringsAsFactors = FALSE
        )
    }

    y_hat_train <- predict(model, newdata = x_train)
    y_hat_test <- predict(model, newdata = x_test)

    # Performance metrics calculation
    mae_train <- mean(abs(y_train - y_hat_train))
    mae_test <- mean(abs(y_test - y_hat_test))

    sse_train <- sum((y_train - y_hat_train)^2)
    sse_test <- sum((y_test - y_hat_test)^2)

    tss_train <- sum((y_train - mean(y_train))^2)
    tss_test <- sum((y_test - mean(y_test))^2)

    rsq_train <- 1 - (sse_train / tss_train)
    rsq_test <- 1 - (sse_test / tss_test)

    rmse_train <- sqrt(mean((y_train - y_hat_train)^2))
    rmse_test <- sqrt(mean((y_test - y_hat_test)^2))

    # Appending results to the dataframe
    new_row <- data.frame(
        Model = model_name,
        SSE_train = sse_train,
        SSE_test = sse_test,
        R_squared_train = rsq_train,
        R_squared_test = rsq_test,
        RMSE_train = rmse_train,
        RMSE_test = rmse_test,
        MAE_train = mae_train,
        MAE_test = mae_test
    )

    df_results <- rbind(df_results, new_row)

    # Returning the updated dataframe
    return(df_results)
}


#' Display Model Results using Datatable
#'
#' This function displays a dataframe containing model performance metrics using the `datatable` function.
#'
#' @param df_results A dataframe containing model performance metrics.
#' @param caption Optional caption for the table (default: "Model Comparison").
#' @return NULL (it displays the table but doesn't return a value).

view_model_results <- function(df_results, caption) {
  # Identifying numeric columns other than "Model", "R_squared_train", and "R_squared_test"
  cols_to_round <- setdiff(names(df_results[sapply(df_results, is.numeric)]), c("Model", "R_squared_train", "R_squared_test"))

  # Round these specific columns to 2 decimal places
  df_results[cols_to_round] <- lapply(df_results[cols_to_round], round, 2)

  # Round R-squared columns to 4 decimal places
  df_results$R_squared_train <- round(df_results$R_squared_train, 5)
  df_results$R_squared_test <- round(df_results$R_squared_test, 5)

  # Display the dataframe using datatable
  datatable(
    df_results,
    caption = caption,
    options = list(
      paging = FALSE,
      autoWidth = TRUE,
      scrollX = TRUE,
      fixedColumns = list(leftColumns = 1)
    )
  )
}


extract_and_display_features <- function(model, full_df, target_var) {
  features_used <- setdiff(names(coef(model)), "(Intercept)")
  features_used <- features_used[features_used != "(Intercept)"]

  # Get all the feature names from the full dataset, excluding the target variable
  all_features <- names(full_df)
  all_features <- all_features[all_features != target_var]

  # Identify features not used in the model and print if desired
  # unused_features <- setdiff(all_features, features_used)

  # Return the features that were used as a character vector
  return(features_used)
}

# Function to create a linear model based on a subset of features
create_model <- function(df, target_var, features) {
  # Construct the formula for lm()
  formula <- as.formula(paste(target_var, "~", paste(features, collapse = "+")))
  # Fit the linear model
  return(lm(formula, data = df))
}

prepare_data <- function(model, train_df, test_df, target_var) {
  # Extracting used features from the model
  used_features <- extract_and_display_features(model, train_df, target_var = target_var)

  # Create x_train and x_test without the target variable
  x_train <- subset(train_df, select = used_features)
  x_test <- subset(test_df, select = used_features)

  y_train <- train_df[[target_var]]
  y_test <- test_df[[target_var]]

  return(list(x_train = x_train, x_test = x_test, y_train = y_train, y_test = y_test))
}

evaluate_model <- function(model_name, model, train_df, test_df, target_var, df_results) {
  # Filter data into x and y subsets
  data <- prepare_data(model, train_df, test_df, target_var = target_var)

  # Adding model performance to the results dataframe
  df_results <- add_model_performance(
    model_name = model_name,
    model = model,
    x_train = data$x_train,
    y_train = data$y_train,
    x_test = data$x_test,
    y_test = data$y_test,
    df_results = df_results  # Or pass existing df_results if available
  )

  return(df_results)
}

# Create an empty coefficients data frame if it doesn't exist, otherwise add rows to it
create_coefficients_df <- function(model, model_name = "Your Model Name", coefficients_df = NULL) {
  # Check the type of model and extract coefficients accordingly
  if (inherits(model, "lm")) {
    coefficients <- coef(model)
  } else if (inherits(model, "dgCMatrix")) { # If it's a matrix from glmnet
    coefficients <- as.vector(model) # Convert matrix to vector
    names(coefficients) <- rownames(model) # Use rownames from matrix as names
  } else {
    stop("Input model is not a recognized type.")
  }

  # Prepare the model data as a dataframe row
  model_row <- data.frame(t(coefficients), check.names = FALSE)
  model_row$Model_Name <- model_name  # Add the model name as a new column

  # If coefficients_df is not provided, initialize a new data frame
  if (is.null(coefficients_df)) {
    coefficients_df <- model_row
  } else {
    # Align the model_row with the coefficients_df
    # Adding NA columns for missing features in coefficients_df
    missing_cols_in_df <- setdiff(names(model_row), names(coefficients_df))
    if (length(missing_cols_in_df) > 0) {
      coefficients_df[missing_cols_in_df] <- NA
    }

    # Adding NA columns for missing features in model_row
    missing_cols_in_row <- setdiff(names(coefficients_df), names(model_row))
    if (length(missing_cols_in_row) > 0) {
      model_row[missing_cols_in_row] <- NA
    }

    # Ensure both dataframes have the same column order
    model_row <- model_row[names(coefficients_df)]

    # Bind the new row to the existing dataframe
    coefficients_df <- rbind(coefficients_df, model_row)
  }

  # Sort the columns alphabetically, except 'Model_Name'
  cols_order <- c("Model_Name", sort(setdiff(names(coefficients_df), "Model_Name")))
  coefficients_df <- coefficients_df[, cols_order]

  return(coefficients_df)
}
```

### 4.2 Model Building

```{r build_inital_linear_model}
# Fit a linear regression model to the training data
linear_model_initial <- lm(price ~ ., data = train_df_linear)

# Initalize and start a coefficients_df to examine later
coefficients_df <- create_coefficients_df(linear_model_initial, "Initial OLS Model")

# Evaluate OLS_linear
df_results <- evaluate_model("OLS_linear", linear_model_initial, train_df_linear, test_df_linear, target_var = 'price', NULL)

# Add results to the df_results to view and sort later
view_model_results(df_results, caption = generate_table_caption("OLS Linear Model Table", section = 4))
```

### 4.3 Model Evaluation

```{r linear_model_evaluation, echo = TRUE, include = TRUE}
# Show inital linear model results
summary(linear_model_initial)
```

#### 4.3.1 Overview of Model Performance

In this section, we provide an in-depth evaluation of the initial linear regression model's performance from a data modeling perspective in the context of predicting house prices.

#### 4.3.2 Model Fit

The model exhibits a respectable fit to the data with an adjusted R-squared value of 0.8174. This metric suggests that approximately 81.74% of the variance in house prices is accounted for by the independent variables included in the model. While this is a strong start, it's important to explore whether there is room for model improvement.

#### 4.3.3 Significant Predictors

Several predictors stand out as statistically significant contributors to the model's predictive power. Notably, variables such as `bedrooms`, `bathrooms`, `sqft_living`, `yr_renovated`, `lat`, `long`, and `view` have coefficients with p-values less than 0.05. These variables have a substantial influence on predicting house prices and are essential components of the model.

#### 4.3.4 Baseline and Intercept

The `(Intercept)` term represents the baseline house price when all other predictors are set to zero. It is crucial in understanding the inherent value of a house. Interpretation of the baseline price helps assess the model's ability to capture the impact of other variables.

#### 4.3.5 Challenges and Missing Coefficients

The model also reveals challenges, such as missing coefficients for some variables (e.g., `sqft_basement`), indicating potential data quality issues. Addressing these gaps is vital to enhance the model's performance and interpretability.

#### 4.3.6 Residual Analysis

Residual analysis indicates that the model's residuals range from -1,415,466 to 4,245,843, suggesting the presence of heteroscedasticity or outliers. Deeper investigation into these issues is necessary to refine the model and ensure its robustness.

#### 4.3.7 Feature Engineering

The model incorporates a wide range of features, including property characteristics, location-based attributes, and temporal variables. These features collectively contribute to its predictive capacity. Ongoing feature engineering efforts should focus on selecting relevant variables and transforming them effectively to improve the model's accuracy.

#### 4.3.8 Model Optimization

While the initial model provides valuable insights, further optimization is warranted. Exploring alternative regression techniques, addressing multicollinearity among predictors, and employing feature selection methods can help enhance model performance.

#### 4.3.9 Next Steps

We will commence with a comprehensive exploration of our dataset in Section 5, utilizing various visualization techniques to gain insights into the relationships between variables. This exploration includes scatter plots (Section 5.1), a correlation matrix (Section 5.2), and in-depth relationship analysis (Section 5.3).

In Section 6, we will employ a stepwise model selection approach to identify the most relevant predictors for our linear regression model. This process will be detailed in Section 6.1. Subsequently, we will present the results for the best linear model iterations in Sections 6.2 and 6.3, followed by a model comparison in Section 6.4.

Ensuring the validity of linear regression assumptions is vital, and we will conduct thorough checks in Section 7. This includes assessing the linearity assumption (Section 7.1), examining the normality of residuals (Section 7.2), and verifying homoscedasticity (constant variance) (Section 7.3).

In Section 8, we will address issues related to heteroscedasticity and multicollinearity. Initially, we will detect and analyze heteroscedasticity in Section 8.1, followed by the presentation of remedial measures in Section 8.2. Subsequently, Section 8.3 will focus on the detection of multicollinearity, with Section 8.4 offering solutions to mitigate its impact.

As we progress, we will explore alternative modeling techniques in Section 9. This includes the implementation of a regression tree model (Section 9.1), a neural network model (Section 9.2), and a support vector machine (SVM) model (Section 9.3). If applicable, we will also consider a logistic regression model (Section 9.4). These alternative models will provide valuable insights and potential enhancements to our predictive capabilities.

```{r reset_section4_counters, echo = FALSE, include = FALSE}
# End of section so set both table and figure to 0
fig_counter <- 0
table_counter <- 0
```

---

## 5. Data Exploration and Visualization
### 5.1 Continuous Variable Plots

#### 5.1.1 Price vs. Square Footage of Living Space

```{r price_vs_sqft_living, echo = TRUE, include = TRUE}
# Scatter plot of Price vs. Square Footage of Living Space
ggplot(data = train_df_non_linear, aes(x = sqft_living, y = price)) +
  geom_point(pch = 20, col = "blue") +
  labs(title = "Price vs. Square Footage of Living Space",
       subtitle = "Seattle Housing Data",
       x = "Sqft Living Space",
       y = "Price",
       caption = generate_figure_caption("Price vs. Square Footage of Living Space", section = 5))
```

In the scatter plot above, we compare the `price` of homes against their `sqft_living` (square footage of interior living space). This visualization allows us to explore the relationship between these two variables.

```{r distribution_sqft_living, echo = TRUE, include = TRUE}
# Distribution of Square Footage of Living Space
ggplot(data = train_df_non_linear, aes(x = sqft_living)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Sqft Living Space",
       x = "Sqft Living Space",
       y = "Density",
       caption = generate_figure_caption("Distribution of Sqft Living Space", section = 5))
```

The histogram above displays the distribution of `sqft_living`. It reveals that the variable is right-skewed, with most homes having smaller living spaces and relatively fewer very large living spaces.

#### 5.1.2 Price vs. Square Footage of Lot

```{r price_vs_sqft_lot, echo = TRUE, include = TRUE}
# Scatter plot of Price vs. Square Footage of Lot
ggplot(data = train_df_non_linear, aes(x = sqft_lot, y = price)) +
  geom_point(pch = 20, col = "green") +
  labs(title = "Price vs. Sqft Lot",
       x = "Sqft Lot",
       y = "Price",
       caption = generate_figure_caption("Price vs. Sqft Lot", section = 5))
```

The scatter plot above compares `price` against `sqft_lot` (square footage of land space). It helps us understand if there's any relationship between the size of the lot and the sale price.

```{r distribution_sqft_lot, echo = TRUE, include = TRUE}
# Distribution of Square Footage of Lot
ggplot(data = train_df_non_linear, aes(x = sqft_lot)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Sqft Lot",
       x = "Sqft Lot",
       y = "Density",
       caption = generate_figure_caption("Distribution of Sqft Lot", section = 5))
```

The histogram above visualizes the distribution of `sqft_lot`. Similar to `sqft_living`, this variable is right-skewed, with most homes having smaller lot sizes and relatively fewer very large lots.

#### 5.1.3 Price vs. Square Footage Above Ground

```{r price_vs_sqft_above, echo = TRUE, include = TRUE}
# Scatter plot of Price vs. Square Footage Above Ground
ggplot(data = train_df_non_linear, aes(x = sqft_above, y = price)) +
  geom_point(pch = 20, col = "red") +
  labs(title = "Price vs. Sqft Above Ground",
       x = "Sqft Above Ground",
       y = "Price",
       caption = generate_figure_caption("Price vs. Sqft Above Ground", section = 5))
```

In the scatter plot above, we compare `price` against `sqft_above` (square footage of the interior housing space above ground level). This analysis helps us explore the impact of above-ground living space on home prices.

```{r distribution_sqft_above, echo = TRUE, include = TRUE}
# Distribution of Square Footage Above Ground
ggplot(data = train_df_non_linear, aes(x = sqft_above)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Sqft Above Ground",
       x = "Sqft Above Ground",
       y = "Density",
       caption = generate_figure_caption("Distribution of Sqft Above Ground", section = 5))
```

The histogram above shows the distribution of `sqft_above`. It suggests that most homes have similar above-ground square footage, with relatively fewer having significantly larger or smaller above-ground spaces.

#### 5.1.4 Price vs. Square Footage of Basement

Excluding homes that do not have a basement.

```{r price_vs_sqft_basement, echo = TRUE, include = TRUE}
# Scatter plot of Price vs. Square Footage of Basement (excluding 0 sqft basement)
# Filter data for non-zero sqft_basement
filtered_data <- train_df_non_linear[train_df_non_linear$sqft_basement > 0,]

# Scatter plot with custom caption
ggplot(data = filtered_data, aes(x = sqft_basement, y = price)) +
  geom_point(pch = 20, col = "purple") +
  labs(title = "Price vs. Sqft Basement",
       x = "Sqft Basement",
       y = "Price",
       caption = generate_figure_caption("Price vs. Sqft Basement (Non-Zero Values)", section = 5))
```

The scatter plot above compares `price` against `sqft_basement` (square footage of the interior housing space below ground level). This visualization helps us understand if the presence and size of a basement influence home prices.

```{r distribution_sqft_basement, echo = TRUE, include = TRUE}
# Distribution of Square Footage of Basement (excluding 0 values)
# Histogram with custom caption
ggplot(data = filtered_data, aes(x = sqft_basement)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Sqft Basement",
       x = "Sqft Basement",
       y = "Density",
       caption = generate_figure_caption("Distribution of Sqft Basement (Non-Zero Values)", section = 5))
```

The histogram above visualizes the distribution of `sqft_basement`. It indicates that most homes have little to no basement space, while some have larger basement areas.

#### 5.1.5 Price vs. Year Built

```{r price_vs_yr_built, echo = TRUE, include = TRUE}
# Scatter plot of Price vs. Year Built
ggplot(data = train_df_non_linear, aes(x = yr_built, y = price)) +
  geom_point(pch = 20, col = "orange") +
  labs(title = "Price vs. Year Built",
       x = "Year Built",
       y = "Price",
       caption = generate_figure_caption("Price vs. Year Built", section = 5))
```

The scatter plot above compares `price` against the year when homes were initially built (`yr_built`). This analysis helps us understand how the age of a home relates to its sale price.

```{r distribution_yr_built, echo = TRUE, include = TRUE}
# Distribution of Year Built
ggplot(data = train_df_non_linear, aes(x = yr_built)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Year Built",
       x = "Year Built",
       y = "Density",
       caption = generate_figure_caption("Distribution of Year Built", section = 5))
```

The histogram above displays the distribution of `yr_built`. It provides insights into the distribution of home ages in the dataset.

#### 5.1.6 Price vs. Year of Last Renovation

Excluding homes that did not have a documented renovation.

```{r price_vs_yr_renovated, echo = TRUE, include = TRUE}
# Find lowest non-zero year renovated
lowest_non_zero_renovation_year <- min(train_df_non_linear$yr_renovated[train_df_non_linear$yr_renovated > 0]) - 1

# Filter data for non-zero yr_renovated
filtered_data <- train_df_non_linear[train_df_non_linear$yr_renovated > 0,]

# Scatter plot of Price vs. Year Renovated
lowest_non_zero_renovation_year <- min(train_df_non_linear$yr_renovated[train_df_non_linear$yr_renovated > 0]) - 1
ggplot(data = filtered_data, aes(x = yr_renovated, y = price)) +
  geom_point(pch = 20, col = "brown") +
  labs(title = "Price vs. Year Renovated",
       x = "Year Renovated",
       y = "Price",
       caption = generate_figure_caption("Price vs. Year Renovated (Non-Zero Values)", section = 5)) +
  xlim(c(lowest_non_zero_renovation_year, max(train_df_non_linear$yr_renovated)))
```

In the scatter plot above, we compare `price` against the year of the last renovation (`yr_renovated`). This analysis helps us understand whether recent renovations impact home prices.

```{r distribution_yr_renovated, echo = TRUE, include = TRUE, warning = FALSE}
# Find lowest non-zero year renovated
lowest_non_zero_renovation_year <- min(train_df_non_linear$yr_renovated[train_df_non_linear$yr_renovated > 0]) - 1

# Filter data for non-zero yr_renovated
filtered_data <- train_df_non_linear[train_df_non_linear$yr_renovated > 0,]

# Histogram of Year Renovated
ggplot(data = filtered_data, aes(x = yr_renovated)) +
  geom_histogram(fill = "orange") +
  labs(title = "Histogram of Year Renovated",
       x = "Year Renovated",
       y = "Density",
       caption = generate_figure_caption("Histogram of Year Renovated (Non-Zero Values)", section = 5)) +
  xlim(c(lowest_non_zero_renovation_year, max(train_df_non_linear$yr_renovated)))
```

The histogram above visualizes the distribution of `yr_renovated`. It provides insights into the distribution of renovation years in the dataset.

#### 5.1.7 Price vs. Distance to Convergence

```{r price_vs_distance_to_convergence, echo = TRUE, include = TRUE}
# Scatter plot of Price vs. Distance to Convergence
ggplot(data = train_df_non_linear, aes(x = distance_to_convergence, y = price)) +
  geom_point(pch = 20, col = "violet") +
  labs(title = "Price vs. Distance to Convergence",
       x = "Distance to Convergence",
       y = "Price",
       caption = generate_figure_caption("Price vs. Distance to Convergence", section = 5))
```

The scatter plot above compares `price` against `distance_to_convergence`. This analysis helps us explore whether the distance to a convergence point impacts home prices.

```{r distribution_distance_to_convergence, echo = TRUE, include = TRUE}
# Distribution of Distance to Convergence
ggplot(data = train_df_non_linear, aes(x = distance_to_convergence)) +
  geom_histogram(bins = 50) +
  labs(title = "Distribution of Distance to Convergence",
       x = "Distance to Convergence",
       y = "Density",
       caption = generate_figure_caption("Distribution of Distance to Convergence", section = 5))
```

### 5.2 Categorical Variable Analysis

The distribution and count of categorical variables such as `bedrooms`, `bathrooms`, `floors`, `waterfront`, `view`, `condition`, and `grade` are analyzed.

#### 5.2.1 Price vs. Bedrooms

```{r price_vs_bedrooms, echo = TRUE, include = TRUE}
# Convert bedrooms to factor
train_df_non_linear$bedrooms_factor <- factor(train_df_non_linear$bedrooms)

# Binned Boxplot of Price vs. Bedrooms
ggplot(data = train_df_non_linear, aes(x = bedrooms_factor, y = price)) +
  geom_boxplot(fill = "blue") +
  labs(title = "Price vs. Bedrooms",
       x = "Bedrooms",
       y = "Price",
       caption = generate_figure_caption("Price vs. Bedrooms", section = 5))
```

The scatter plot above compares `price` against the number of `bedrooms`. This visualization helps us understand how the number of bedrooms influences home prices.

```{r distribution_bedrooms, echo = TRUE, include = TRUE}
# Filter data excluding 33 bedrooms
filtered_bedrooms <- train_df_non_linear$bedrooms[train_df_non_linear$bedrooms != 33]

# Calculate frequencies of each bedroom count
bedroom_frequencies <- table(filtered_bedrooms)

ggplot(data = data.frame(filtered_bedrooms = as.factor(names(bedroom_frequencies)),
                        filtered_counts = as.numeric(bedroom_frequencies)),
       aes(x = filtered_bedrooms, y = filtered_counts)) +
  geom_bar(stat = "identity", fill = "blue") +
  labs(title = "Distribution of Bedrooms (Excluding 33 Bedrooms)",
       x = "Number of Bedrooms",
       y = "Frequency",
       caption = generate_figure_caption("Distribution of Bedrooms (Excluding 33 Bedrooms)", section = 5))
```

The bar plot above displays the distribution of the `bedrooms` variable, showing the frequency of each bedroom count.

#### 5.2.2 Price vs. Bathrooms

```{r price_vs_bathrooms, echo = TRUE, include = TRUE}
# Convert bathrooms to factor
train_df_non_linear$bathrooms_factor <- factor(train_df_non_linear$bathrooms)

# Binned Boxplot of Price vs. Bathrooms
ggplot(data = train_df_non_linear, aes(x = bathrooms_factor, y = price)) +
  geom_boxplot(fill = "green") +
  labs(title = "Price vs. Bathrooms",
       x = "Bathrooms",
       y = "Price",
       caption = generate_figure_caption("Price vs. Bathrooms", section = 5))
```

In the scatter plot above, we compare `price` against the number of `bathrooms`. This analysis helps us explore the relationship between the number of bathrooms and home prices.

```{r distribution_bathrooms, echo = TRUE, include = TRUE}
# Get data for bar plot
bathrooms_counts <- table(train_df_non_linear$bathrooms)
bathrooms <- as.numeric(names(bathrooms_counts))
counts <- as.numeric(bathrooms_counts)

# Bar plot for the distribution of Bathrooms
ggplot(data = data.frame(bathrooms, counts), aes(x = bathrooms, y = counts)) +
  geom_bar(stat = "identity", fill = "green") +
  labs(title = "Distribution of Bathrooms",
       x = "Number of Bathrooms",
       y = "Frequency",
       caption = generate_figure_caption("Distribution of Bathrooms", section = 5))
```

The bar plot above visualizes the distribution of the `bathrooms` variable, showing the frequency of each bathroom count.

#### 5.2.3 Price vs. Floors

```{r price_vs_floors, echo = TRUE, include = TRUE}
# Binned Boxplot of Price vs. Floors
ggplot(data = train_df_non_linear, aes(x = floors, y = price, group = floors)) +
  geom_boxplot(fill = "orange") +
  labs(title = "Price vs. Floors",
       x = "Floors",
       y = "Price",
       caption = generate_figure_caption("Price vs. Floors", section = 5))
```

The scatter plot above compares `price` against the number of `floors`. This analysis helps us understand how the number of floors in a home relates to its sale price.

```{r distribution_floors, echo = TRUE, include = TRUE}
floors_counts <- table(train_df_non_linear$floors)
floors <- as.numeric(names(floors_counts))
counts <- as.numeric(floors_counts)

# Bar plot for the distribution of Floors
ggplot(data = data.frame(floors, counts), aes(x = floors, y = counts)) +
  geom_bar(stat = "identity", fill = "orange") +
  labs(title = "Distribution of Floors",
       x = "Number of Floors",
       y = "Frequency",
       caption = generate_figure_caption("Distribution of Floors", section = 5))
```

The bar plot above displays the distribution of the `floors` variable, showing the frequency of each floor count.

#### 5.2.4 Price vs. Waterfront

```{r price_vs_waterfront, echo = TRUE, include = TRUE}
ggplot(data = train_df_non_linear, aes(x = waterfront_1, y = price, group = waterfront_1)) +
  geom_boxplot(fill = "purple") +
  labs(title = "Price vs. Waterfront",
       x = "Waterfront",
       y = "Price",
       caption = generate_figure_caption("Price vs. Waterfront", section = 5),
       fill = "Waterfront",
       levels = c("No", "Yes"))  # Labels for waterfront status
```

In the scatter plot above, we compare `price` against the `waterfront` variable. This visualization helps us explore how having a waterfront view impacts home prices.

```{r distribution_waterfront, echo = TRUE, include = TRUE}
# Get data for bar plot
waterfront_counts <- table(train_df_non_linear$waterfront_1)
waterfront <- as.numeric(names(waterfront_counts))
counts <- as.numeric(waterfront_counts)

# Bar plot for the distribution of Waterfront
ggplot(data = data.frame(waterfront, counts), aes(x = waterfront, y = counts)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Distribution of Waterfront",
       x = "Waterfront",
       y = "Frequency",
       caption = generate_figure_caption("Distribution of Waterfront", section = 5),
       fill = "Waterfront",
       levels = c("No", "Yes"))  # Labels for waterfront status

```

The bar plot above visualizes the distribution of the `waterfront` variable, showing the frequency of waterfront and non-waterfront properties.

#### 5.2.5 Price vs. View

```{r price_vs_view, echo = TRUE, include = TRUE}
# Convert view categories from dummy variables to a factor for better labeling in ggplot
train_df_non_linear$view_category <- factor(apply(train_df_non_linear[, c("view_0", "view_1", "view_2", "view_3", "view_4")], 1, function(x) which(x == 1)),
                                             labels = c("View 0", "View 1", "View 2", "View 3", "View 4"))

# Create the boxplot with ggplot2
ggplot(train_df_non_linear, aes(x = view_category, y = price)) +
  geom_boxplot(fill = "brown") +
  labs(title = "Price vs. View Quality",
       x = "View Quality",
       y = "Price",
       caption = generate_figure_caption("Boxplot of Price vs. View Quality", section = 5))
```

The scatter plot above compares `price` against the `view` variable, which represents the quality of the property's view. This analysis helps us explore how the view quality impacts home prices.

```{r distribution_view, echo = TRUE, include = TRUE}
# Calculate frequencies of each view quality rating
view_frequencies <- colSums(train_df_non_linear[, c("view_0", "view_1", "view_2", "view_3", "view_4")])

# Convert frequencies to data frame for ggplot2
view_df <- data.frame(View = names(view_frequencies), Frequency = view_frequencies)

# Create the bar plot with ggplot2
ggplot(view_df, aes(x = View, y = Frequency)) +
  geom_bar(stat = "identity", fill = "purple") +
  labs(title = "Distribution of View Quality",
       x = "View Quality",
       y = "Frequency",
       caption = generate_figure_caption("Distribution of View Quality", section = 5))
```

The bar plot above displays the distribution of the `view` variable, showing the frequency of different view quality ratings.

#### 5.2.6 Price vs. Condition

```{r price_vs_condition, echo = TRUE, include = TRUE}
# Convert condition categories from dummy variables to a factor
train_df_non_linear$condition_category <- factor(apply(train_df_non_linear[, c("condition_1", "condition_2", "condition_3", "condition_4", "condition_5")], 1, function(x) which(x == 1)),
                                                 labels = c("Condition 1", "Condition 2", "Condition 3", "Condition 4", "Condition 5"))

# Create the boxplot with ggplot2
ggplot(train_df_non_linear, aes(x = condition_category, y = price)) +
  geom_boxplot(fill = "blue") +
  labs(title = "Price vs. Condition",
       x = "Condition",
       y = "Price",
       caption = generate_figure_caption("Boxplot of Price vs. Condition", section = 5))
```

In the scatter plot above, we compare `price` against the `condition` variable, which represents the condition of the property. This analysis helps us explore how property condition relates to home prices.

```{r distribution_condition, echo = TRUE, include = TRUE}
# Calculate frequencies of each condition rating
condition_frequencies <- colSums(train_df_non_linear[, c("condition_1", "condition_2", "condition_3", "condition_4", "condition_5")])

# Convert frequencies to a data frame for ggplot2
condition_df <- data.frame(Condition = names(condition_frequencies), Frequency = condition_frequencies)

# Create the bar plot with ggplot2
ggplot(condition_df, aes(x = Condition, y = Frequency)) +
  geom_bar(stat = "identity", fill = "green") +
  labs(title = "Distribution of Condition",
       x = "Condition Rating",
       y = "Frequency",
       caption = generate_figure_caption("Distribution of Condition", section = 5))
```

The bar plot above visualizes the distribution of the `condition` variable, showing the frequency of different condition ratings.

#### 5.2.7 Price vs. Grade

```{r price_vs_grade, echo = TRUE, include = TRUE}
# First, identify all grade-related columns in the dataframe
grade_columns <- grep("grade_", names(train_df_non_linear), value = TRUE)

# Convert dummy variables back to a single categorical variable representing the grade
train_df_non_linear$grade_category <- apply(train_df_non_linear[, grade_columns], 1, function(row) {
  if (all(is.na(row))) {
    return(NA)  # Return NA if all values in the row are NA
  } else {
    idx <- which(row == 1, arr.ind = TRUE)
    return(if(length(idx) > 0) idx else NA)  # Return the index of the grade, or NA if none is 1
  }
})

# Extract grade labels from column names, replacing underscores with hyphens for better readability
grade_labels <- sub("grade_", "", grade_columns) # Remove 'grade_' prefix
grade_labels <- gsub("_", "-", grade_labels) # Replace underscores with hyphens

# Create a boxplot of Price vs. Grade
ggplot(train_df_non_linear, aes(x = factor(grade_category, labels = grade_labels), y = price)) +
  geom_boxplot(fill = "green") +
  labs(title = "Price vs. Grade",
       x = "Grade",
       y = "Price",
       caption = generate_figure_caption("Boxplot of Price vs. Grade", section = 5))
```

The scatter plot above compares `price` against the `grade` variable, which has been aggregated into categories as per the provided header. This analysis helps us explore how the grade of construction and design impacts home prices.

```{r distribution_grade, echo = TRUE, include = TRUE}
# Histogram for the Distribution of Grade
# Convert the grade category to a numeric variable for histogram plotting
train_df_non_linear$grade_category_numeric <- as.numeric(train_df_non_linear$grade_category)

# Define breaks for histogram
num_breaks <- length(unique(train_df_non_linear$grade_category_numeric, na.rm = TRUE))
hist_breaks <- seq(min(train_df_non_linear$grade_category_numeric, na.rm = TRUE) - 0.5,
                   max(train_df_non_linear$grade_category_numeric, na.rm = TRUE) + 0.5,
                   length.out = num_breaks + 1)

# Create a histogram with ggplot2
ggplot(train_df_non_linear, aes(x = grade_category_numeric)) +
  geom_histogram(fill = "purple", breaks = hist_breaks) +
  scale_x_continuous(breaks = seq_along(grade_labels), labels = grade_labels) +
  labs(title = "Distribution of Grade",
       x = "Grade",
       y = "Frequency",
       caption = generate_figure_caption("Histogram of Distribution of Grade", section = 5))
```

The bar plot above displays the distribution of the `grade_category` variable, showing the frequency of different grade categories.

### 5.3 Correlation Analysis

Understanding how continuous variables correlate with each other and, more importantly, with the target variable `price`.

#### 5.3.1 Correlation Matrix

```{r correlation_analysis, echo = TRUE, include = TRUE}
# Correlation Matrix of Numeric Variables
cor_matrix <- cor(train_df_non_linear[sapply(train_df_non_linear, is.numeric)])
```

```{r correlation_matrix_table, echo = TRUE, include = TRUE}
# Create a table of sorted correlation values
cor_table <- as.data.frame(sort(cor_matrix[,"price"], decreasing = TRUE))

# Display the top 20 correlation values
top_20_corr <- cor_table[1:20, , drop = FALSE]
```

```{r corr_matrix_table_display, echo = FALSE, include = TRUE, results = 'asis'}
# Specify the height for the scrollable div
table_height <- "400px"

# Create a scrollable div and place the table inside it
cat(sprintf('<div style="overflow-y: scroll; max-height: %s;">', table_height))
kable(top_20_corr, col.names = c("Variable", "Correlation with Price"), caption = generate_table_caption("Top 20 Correlation Values with Price", section = 5))
cat('</div>')
```

#### 5.3.2 Correlation Graphics Analysis

In the tables presented above, we've showcased the top 20 correlation values concerning the target variable, `price`, with the values sorted by their absolute magnitudes. Here are some crucial observations from this analysis:

   1. **Positive Correlations with Price**:
      - Variables like `sqft_living`, `sqft_above`, `sqft_living15`, and `bathrooms` exhibit robust positive correlations with the target variable (`price`). This implies that as these features increase, house prices tend to increase correspondingly.
      - Features such as `grade_11_13`, `view_4`, and `grade_8_10` also demonstrate positive correlations, indicating that properties with higher grades and better views tend to command higher prices.

   2. **Negative Correlations with Price**:
      - There are no negative correlations among the top 20 correlated variables. This suggests that none of the examined features strongly suggest a decrease in house price as they increase.

   3. **Feature Importance**:
      - The strength of these correlations provides insights into the importance of variables in predicting house prices. Variables like `sqft_living` and `grade_11_13` emerge as strong predictors of price.
      - Location-related variables, such as `zipcode_98004`, `zipcode_98039`, and `zipcode_98040`, also exhibit noteworthy positive correlations, underscoring the significance of location in price determination.

#### 5.3.3 Correlation Heatmap

```{r correlation_heatmap, echo = TRUE, include = TRUE}
# Heatmap of the top 20 correlation values
# Filter the top 20 correlation values
top_20_corr_variables <- rownames(top_20_corr)
top_20_corr_matrix <- cor_matrix[top_20_corr_variables, top_20_corr_variables]

# Create a heatmap
ggplot(melt(top_20_corr_matrix), aes(Var1, Var2, fill = value)) +
  geom_tile() +
  labs(title = "Top 20 Correlations",
       x = "Variable",
       y = "Variable",
       caption = generate_figure_caption("Heatmap showing the top 20 correlations", section = 5)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### 5.3.4 Correlation Matrix for Multicollinearity

```{r correlation_matrix_for_multicollinearity, echo = TRUE, include = TRUE, results = 'asis'}
# Selecting predictors and excluding the response variable 'price'
predictors <- dplyr::select(train_df_linear, -price)

# Convert factors to numeric
numeric_predictors <- predictors %>%
  mutate(across(where(is.factor), as.numeric)) %>%
  mutate(across(where(is.character), ~as.numeric(as.factor(.))))

# Calculate the correlation matrix
corr_matrix <- cor(numeric_predictors, use = "pairwise.complete.obs")

# Convert the correlation matrix to a long format
correlated_pairs_df <- melt(corr_matrix)

# Filter out redundant pairs (keep only lower triangle of the matrix)
correlated_pairs_df <- correlated_pairs_df %>%
  filter(Var1 != Var2) %>%  # Remove self-correlations
  filter(abs(value) > 0.8) %>%
  filter(match(Var1, rownames(corr_matrix)) < match(Var2, rownames(corr_matrix)))

# Rename columns for clarity
correlated_pairs_df <- correlated_pairs_df %>%
  rename(Variable1 = Var1, Variable2 = Var2, Correlation = value)

# Output the table of highly correlated pairs using knitr::kable()
knitr_table <- kable(
  correlated_pairs_df,
  caption = generate_table_caption("Highly Correlated Variable Pairs", section = 5),
  format = "markdown"
)
print(knitr_table)
```

#### 5.3.5 Detailed Explanation for Removal

##### 5.3.5.1 `sqft_above` & `sqft_living`

The removal of `sqft_above` and `sqft_living` is justified due to their high correlation coefficient of 0.8744114. `sqft_above` represents the square footage of the living area above ground, while `sqft_living` encompasses the total square footage of living space. Since `sqft_above` is a subset of `sqft_living`, it is likely to contain redundant information, making it less valuable for our model.

##### 5.3.5.2 `month_sold` & `week_of_year`

The variables `month_sold` and `week_of_year` exhibit a remarkably high correlation coefficient of 0.9955447. These variables are intrinsically correlated as they both pertain to the date of the house sale. While `day_of_year` provides the most detailed temporal information, retaining both `week_of_year` and `month_sold` may lead to multicollinearity issues. It's advisable to consider removing one of these variables to mitigate multicollinearity while preserving the most granular date-related information.

##### 5.3.5.3 `month_sold` & `day_of_year`

Similar to the previous case, `month_sold` and `day_of_year` demonstrate a high correlation coefficient of 0.9958281. Both variables are related to the date of the house sale. Given that `day_of_year` provides the most granular temporal information, it may be preferred to retain it while considering the removal of `month_sold` to address multicollinearity concerns.

##### 5.3.5.4 `week_of_year` & `day_of_year`

The correlation coefficient of 0.9996951 between `week_of_year` and `day_of_year` indicates an extremely high correlation. Both variables are associated with the date of sale. Given the granularity of `day_of_year`, retaining it and potentially removing `week_of_year` can be a strategy to reduce multicollinearity while retaining essential date-related information.

##### 5.3.5.5 `condition_4` & `condition_3`

The variables `condition_4` and `condition_3` display a notable negative correlation coefficient of -0.8095157. These variables are derived from the categorical variable indicating the condition of the house. Through one-hot encoding, binary variables were created for each condition. Since these conditions are mutually exclusive, they exhibit a negative correlation. Consideration can be given to keeping one condition as a reference group and discarding the other, or reverting to using the original categorical variable to effectively capture overall house condition.

##### 5.3.5.6 `grade_Above_Average` & `grade_Average`

The correlation coefficient of -0.9954905 between `grade_Above_Average` and `grade_Average` highlights a strong negative correlation. These variables represent different grade categories of houses. Such a high correlation suggests that retaining both variables may introduce multicollinearity into the model. Decisions can be made to keep one of these variables as a representative of house grade or explore alternative encoding strategies.

By addressing the removal of these highly correlated variable pairs, our primary goal is to mitigate multicollinearity issues. Multicollinearity can distort regression coefficient estimates, inflate standard errors, and potentially obscure the statistical significance of predictors. The objective is to retain variables that provide unique and informative contributions to the model's prediction of house prices.

### 5.5 Temporal Trends Analysis

Analyzing the influence of time-related features such as `month`, and `season` on house prices.

#### 5.5.1 Monthly Trends in Average House Prices

```{r monthly_trends_in_average_house_prices, echo = TRUE, include = TRUE}
# Monthly Trends in Average House Prices
monthly_trends <- aggregate(train_df_non_linear$price, by = list(train_df_non_linear$month_sold), FUN = mean)
colnames(monthly_trends) <- c("Month", "Average_Price")

# Find the global maximum
global_max <- monthly_trends[which.max(monthly_trends$Average_Price), ]

ggplot(monthly_trends, aes(x = Month, y = Average_Price)) +
  geom_line() +
  geom_text(data = global_max, aes(x = Month, y = Average_Price, label = paste("Global Max:", round(Average_Price, 2))), vjust = -0.5) +
  labs(title = "Monthly Trends in Average House Prices", x = "Month", y = "Average Price", caption = generate_figure_caption("Monthly Trends in Average House Prices", section = 5)) +
  scale_x_continuous(breaks = 1:12, labels = month.name) +
  theme_minimal()
```

```{r monthly_trends_count, echo = TRUE, include = TRUE}
# Monthly Trends in Count of Homes Sold
monthly_counts <- table(train_df_non_linear$month_sold)
months <- factor(1:12, labels = month.name)
monthly_counts_df <- data.frame(Month = months, Count = as.numeric(monthly_counts))

# Find the global maximum
global_max_count <- monthly_counts_df[which.max(monthly_counts_df$Count), ]

ggplot(monthly_counts_df, aes(x = Month, y = Count, group = 1)) +
  geom_bar(stat = "identity") +
  geom_text(data = global_max_count, aes(x = Month, y = Count, label = paste("Global Max:", Count)), vjust = -0.5) +  # Add label for global maximum
  labs(title = "Monthly Trends in Count of Homes Sold", x = "Month", y = "Count of Homes Sold", caption = generate_figure_caption("Monthly Trends in Count of Homes Sold", section = 5)) +
  theme_minimal()
```

#### 5.5.2 Seasonal Trends in Average House Prices

```{r seasonal_trends_in_average_house_prices, echo = TRUE, include = TRUE}
# Aggregate average price for each season
seasonal_trends <- data.frame(
  Season = c("Winter", "Spring", "Summer", "Fall"),
  Average_Price = c(
    mean(train_df_non_linear$price[train_df_non_linear$season_Winter == 1]),
    mean(train_df_non_linear$price[train_df_non_linear$season_Spring == 1]),
    mean(train_df_non_linear$price[train_df_non_linear$season_Summer == 1]),
    mean(train_df_non_linear$price[train_df_non_linear$season_Fall == 1])
  )
)

# Find the global maximum
global_max_seasonal <- seasonal_trends[which.max(seasonal_trends$Average_Price), ]

# Plotting
ggplot(seasonal_trends, aes(x = Season, y = Average_Price, fill = Season)) +
  geom_bar(stat = "identity") +
  geom_text(data = global_max_seasonal, aes(label = paste("Global Max:", round(Average_Price, 2)), y = Average_Price), vjust = -0.5) +
  labs(title = "Seasonal Trends in Average House Prices", x = "Season", y = "Average Price", caption = generate_figure_caption("Seasonal Trends in Average House Prices", section = 5)) +
  theme_minimal()

```

```{r seasonal_trends_count, echo = TRUE, include = TRUE}
# Count homes sold for each season
seasonal_counts <- c(
  sum(train_df_non_linear$season_Winter == 1),
  sum(train_df_non_linear$season_Spring == 1),
  sum(train_df_non_linear$season_Summer == 1),
  sum(train_df_non_linear$season_Fall == 1)
)
seasonal_counts_df <- data.frame(Season = c("Winter", "Spring", "Summer", "Fall"), Count = seasonal_counts)

# Find the global maximum
global_max_count_seasonal <- seasonal_counts_df[which.max(seasonal_counts_df$Count), ]

# Plotting
ggplot(seasonal_counts_df, aes(x = Season, y = Count, fill = Season)) +
  geom_bar(stat = "identity") +
  geom_text(data = global_max_count_seasonal, aes(label = paste("Global Max:", Count), y = Count), vjust = -0.5) +
  labs(title = "Seasonal Trends in Count of Homes Sold", x = "Season", y = "Count of Homes Sold", caption = generate_figure_caption("Seasonal Trends in Count of Homes Sold", section = 5)) +
  theme_minimal()

```

#### 5.5.3 Week of the Year Trends in Average House Prices

```{r week_of_the_year_trends_in_average_house_prices, echo = TRUE, include = TRUE}
# Week of the Year Trends in Average House Prices
weekly_trends <- aggregate(train_df_non_linear$price, by = list(train_df_non_linear$week_of_year), FUN = mean)
colnames(weekly_trends) <- c("Week_of_Year", "Average_Price")

# Find the global maximum
global_max_weekly <- weekly_trends[which.max(weekly_trends$Average_Price), ]

ggplot(weekly_trends, aes(x = Week_of_Year, y = Average_Price)) +
  geom_line() +
  geom_text(data = global_max_weekly, aes(x = Week_of_Year, y = Average_Price, label = paste("Global Max:", round(Average_Price, 2))), vjust = -0.5) +  # Add label for global maximum
  labs(title = "Weekly Trends in Average House Prices", x = "Week of Year", y = "Average Price", caption = generate_figure_caption("Weekly Trends in Average House Prices", section = 5)) +
  theme_minimal()
```

```{r week_of_the_year_trends_count, echo = TRUE, include = TRUE}
# Weekly Trends in Count of Homes Sold
weekly_counts <- table(train_df_non_linear$week_of_year)
weekly_counts_df <- data.frame(Week_of_Year = as.numeric(names(weekly_counts)), Count = as.numeric(weekly_counts))

# Find the global maximum
global_max_count_weekly <- weekly_counts_df[which.max(weekly_counts_df$Count), ]

ggplot(weekly_counts_df, aes(x = Week_of_Year, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(data = global_max_count_weekly, aes(x = Week_of_Year, y = Count, label = paste("Global Max:", Count)), vjust = -0.5) +  # Add label for global maximum
  labs(title = "Weekly Trends in Count of Homes Sold", x = "Week of Year", y = "Count of Homes Sold", caption = generate_figure_caption("Weekly Trends in Count of Homes Sold", section = 5)) +
  theme_minimal()
```

#### 5.5.4 Day of the Year Trends in Average House Prices

```{r day_of_the_year_trends_in_average_house_prices, echo = TRUE, include = TRUE}
# Day of the Year Trends in Average House Prices
daily_trends <- aggregate(train_df_non_linear$price, by = list(train_df_non_linear$day_of_year), FUN = mean)
colnames(daily_trends) <- c("Day_of_Year", "Average_Price")

# Find the global maximum
global_max_daily <- daily_trends[which.max(daily_trends$Average_Price), ]

ggplot(daily_trends, aes(x = Day_of_Year, y = Average_Price)) +
  geom_line() +
  geom_text(data = global_max_daily

, aes(x = Day_of_Year, y = Average_Price, label = paste("Global Max:", round(Average_Price, 2))), vjust = -0.5) +  # Add label for global maximum
  labs(title = "Daily Trends in Average House Prices", x = "Day of Year", y = "Average Price", caption = generate_figure_caption("Daily Trends in Average House Prices", section = 5)) +
  theme_minimal()
```

```{r day_of_the_year_trends_count, echo = TRUE, include = TRUE}
# Daily Trends in Count of Homes Sold
daily_counts <- table(train_df_non_linear$day_of_year)
daily_counts_df <- data.frame(Day_of_Year = as.numeric(names(daily_counts)), Count = as.numeric(daily_counts))

# Find the global maximum
global_max_count_daily <- daily_counts_df[which.max(daily_counts_df$Count), ]

ggplot(daily_counts_df, aes(x = Day_of_Year, y = Count)) +
  geom_bar(stat = "identity") +
  geom_text(data = global_max_count_daily, aes(x = Day_of_Year, y = Count, label = paste("Global Max:", Count)), vjust = -0.5) +  # Add label for global maximum
  labs(title = "Daily Trends in Count of Homes Sold", x = "Day of Year", y = "Count of Homes Sold", caption = generate_figure_caption("Daily Trends in Count of Homes Sold", section = 5)) +
  theme_minimal()
```

- The "Daily Trends in Average House Prices" line chart showcases the average house prices for each day of the year. It helps identify daily patterns and potential price variations that could be influenced by specific dates or events.

### 5.6 Geographical Influence Analysis

Investigating the spatial aspect by analyzing the `distance_to_convergence` variable.

#### 5.6.1 Distance to Convergence Point Map

```{r geographical_price_analysis, echo = TRUE, include = TRUE, results = 'asis'}
# Calculate z-scores for the prices
train_df_non_linear <- train_df_non_linear %>%
  mutate(z_score = scale(price))

# Define z-score intervals and corresponding colors
z_score_intervals <- seq(-3, 3, by = 1)  # Create a sequence of z-scores from -3 to 3
color_sequence <- c("green", "#8fd744", "#fde725", "#f76818ff", "#d7301fff", "#440154")  # From green to dark color

# Calculate price at each z-score interval
price_at_intervals <- sapply(z_score_intervals, function(z) {
  mean(train_df_non_linear$price) + z * sd(train_df_non_linear$price)
})

# Ensure breaks are in ascending order and rounded to the nearest 25k
breaks <- sort(round(price_at_intervals / 25000) * 25000)
breaks <- c(min(train_df_non_linear$price, na.rm = TRUE), breaks, max(train_df_non_linear$price, na.rm = TRUE))

# If there are negative values or values that don't make sense, remove them
breaks <- breaks[breaks >= 0]

# Create color palette with a color for each interval
color_palette <- colorBin(color_sequence, domain = train_df_non_linear$price, bins = breaks, na.color = "#808080")

# Initialize the leaflet map with updated color palette
m <- leaflet(train_df_non_linear) %>%
  addTiles() %>%
  addCircleMarkers(
    lat = ~lat, lng = ~long,
    color = ~color_palette(price),
    fillColor = ~color_palette(price),
    fillOpacity = 0.8,
    radius = 1,  # Small dots
    popup = ~paste("Price: $", formatC(price, format = "f", big.mark = ","), "<br>", "Z-Score: ", round(z_score, 2))
  )

# Define the maximum distance for the distance bands
max_distance <- max(train_df_non_linear$distance_to_convergence, na.rm = TRUE)

# Add distance bands to the map
for (i in seq(2, max_distance, by = 2)) {
  m <- addCircles(m, lat = convergence_point[1], lng = convergence_point[2], radius = i * 1000,
                  color = "grey", weight = 1, fill = FALSE, dashArray = "5, 5")
}

# Add legend and finalize the map
m <- m %>%
  addLegend(
    position = "bottomright",
    pal = color_palette,
    values = ~price,
    title = "Price",
    labFormat = labelFormat(prefix = "$"),
    opacity = 1
  ) %>%
  setView(lng = convergence_point[2], lat = convergence_point[1], zoom = 10)

cat(generate_figure_caption('Distance to Convergence Map', section = 5))
```

```{r display_geo_map, echo = FALSE, include = TRUE}
# Print the map
m
```

#### 5.6.2 Conclusion
This detailed review of the King County house sales dataset underscores the thorough preparation undertaken for the predictive analysis. The dataset's diverse variables, both continuous and categorical, have been meticulously processed and analyzed, providing a robust foundation for developing the predictive model. With the comprehensive EDA and graphical analysis, we gain valuable insights into the correlations and distributions within the data, setting the stage for effective model building and accurate house price prediction.

### 5.7 Removal of Plot Features, Correlation, Multicollinearity and NA Values

```{r remove_plot_features, echo = TRUE, include = TRUE}
# Drop columns created for visualizations in prior steps, columns that have high correlation, multicollinearity or NA values in the model
train_df_non_linear <- train_df_non_linear[, !colnames(train_df_non_linear) %in% c("view_category", "condition_category", "grade_category", "grade_category_numeric", "z_score", "lat", "long", 'sqft_above', 'month_sold', 'week_of_year', 'condition_3', "grade_Below_Average", "bedrooms_factor", "bathrooms_factor")]
train_df_linear <- train_df_linear[, !colnames(train_df_linear) %in% c("view_category", "condition_category", "grade_category", "grade_category_numeric", "z_score", "lat", "long", 'sqft_above', 'month_sold', 'week_of_year', 'condition_3', "grade_Below_Average", "bedrooms_factor", "bathrooms_factor")]
test_df_linear <- test_df_linear[, !colnames(test_df_linear) %in% c("view_category", "condition_category", "grade_category", "grade_category_numeric", "z_score", "lat", "long", 'sqft_above', 'month_sold', 'week_of_year', 'condition_3', "grade_Below_Average", "bedrooms_factor", "bathrooms_factor")]
test_df_non_linear <- test_df_non_linear[, !colnames(test_df_non_linear) %in% c("view_category", "condition_category", "grade_category", "grade_category_numeric", "z_score", "lat", "long", 'sqft_above', 'month_sold', 'week_of_year', 'condition_3', "grade_Below_Average", "bedrooms_factor", "bathrooms_factor")]

# Rebuild linear regression model before performing stepwise
linear_model_initial <- lm(price ~ ., data = train_df_linear)

# Add new coefficients to dataframe
coefficients_df <- create_coefficients_df(
  model = linear_model_initial,
  model_name = "OLS w/o corr",
  coefficients_df = coefficients_df
)
```


```{r reset_section5_counters, echo = FALSE, include = FALSE}
# End of section so set both table and figure to 0
fig_counter <- 0
table_counter <- 0
```

---

## 6. Stepwise Model Selection

### 6.1 Stepwise Methodology


```{r step_model, echo = TRUE, include = TRUE}
# Define a function to fetch dropped features from models and create a kable table
get_dropped_features_table <- function(model_backward, model_forward, best_linear_model, train_df_linear) {
  # Create a data frame to store the results
  results_df <- data.frame(Stepwise_Method = character(), Dropped_Features = character(), stringsAsFactors = FALSE)

  # All possible features (excluding the response variable 'price')
  all_possible_features <- setdiff(colnames(train_df_linear), "price")

  # Function to find the dropped features for a given model
  get_dropped_features <- function(model, all_features) {
    included_features <- names(coef(model))
    dropped_features <- setdiff(all_features, included_features)
    return(paste(dropped_features, collapse = ", "))
  }

  # Get the dropped features for each model
  dropped_backward <- get_dropped_features(model_backward, all_possible_features)
  dropped_forward <- get_dropped_features(model_forward, all_possible_features)
  dropped_both <- get_dropped_features(best_linear_model, all_possible_features)

  # Add results to the data frame
  results_df <- rbind(results_df, data.frame(Stepwise_Method = "OLS_Step_Backward", Dropped_Features = dropped_backward))
  results_df <- rbind(results_df, data.frame(Stepwise_Method = "OLS_Step_Forward", Dropped_Features = dropped_forward))
  results_df <- rbind(results_df, data.frame(Stepwise_Method = "OLS_Step_Both", Dropped_Features = dropped_both))

  # Create the kable table
  table <- kable(
    results_df,
    format = "html",
    caption = generate_table_caption("Dropped Features", section = 6),
    col.names = c("Stepwise Method", "Dropped Features")
  ) %>%
    kable_styling(full_width = TRUE)

  return(table)
}

# Conditional logic based on the update_model_parameters flag
if (update_model_parameters) {
  # Run each of the stepwise regression models and update the JSON file
  # Step model both
  best_linear_model <- ols_step_both_p(linear_model_initial, pent=0.35, prem=0.05)
  features_both <- setdiff(names(coef(best_linear_model$model)), "(Intercept)")

  # Create a formula with the selected features
  formula <- reformulate(features_forward, response = "price")

  # Refit the model with the selected features
  best_linear_model <- lm(formula, data = train_df_linear)

  # Update JSON
  update_model_json("OLS_Step_Both", features_both, json_filepath)

  # Step model Backward
  model_backward <- ols_step_backward_p(linear_model_initial, pent=0.35, prem = 0.05)
  features_backward <- setdiff(names(coef(model_backward$model)), "(Intercept)")

  # Create a formula with the selected features
  formula <- reformulate(features_forward, response = "price")

  # Refit the model with the selected features
  model_backward <- lm(formula, data = train_df_linear)

  # Update JSON
  update_model_json("OLS_Step_Backward", features_backward, json_filepath)

  # Step model foward
  model_forward <- ols_step_forward_p(linear_model_initial, pent=0.35, penter = 0.05)
  features_forward <- setdiff(names(coef(model_forward$model)), "(Intercept)")

  # Create a formula with the selected features
  formula <- reformulate(features_forward, response = "price")

  # Refit the model with the selected features
  model_forward <- lm(formula, data = train_df_linear)

  # Update JSON
  update_model_json("OLS_Step_Forward", features_forward, json_filepath)

  # Load model parameters from JSON and build models
  model_params <- fromJSON(json_filepath)
  # List of features for each model from model_params
  features_both <- model_params$OLS_Step_Both
  features_backward <- model_params$OLS_Step_Backward
  features_forward <- model_params$OLS_Step_Forward

  # Create a temporary dataframe for each model with the selected features
  train_df_both <- train_df_linear[, c("price", features_both)]
  test_df_both <- test_df_linear[, c("price", features_both)]

  train_df_backward <- train_df_linear[, c("price", features_backward)]
  test_df_backward <- test_df_linear[, c("price", features_backward)]

  train_df_forward <- train_df_linear[, c("price", features_forward)]
  test_df_forward <- test_df_linear[, c("price", features_forward)]

  # Fit the linear models using the temporary dataframes
  best_linear_model <- lm(price ~ ., data = train_df_both)
  model_backward <- lm(price ~ ., data = train_df_backward)
  model_forward <- lm(price ~ ., data = train_df_forward)

} else {
  # Load model parameters from JSON and build models
  model_params <- fromJSON(json_filepath)

  # Create models based on the loaded features
  if (all(c("OLS_Step_Both", "OLS_Step_Backward", "OLS_Step_Forward") %in% names(model_params))) {
    # For each model type, create the model using the features stored in the JSON
    # List of features for each model from model_params
    features_both <- model_params$OLS_Step_Both
    features_backward <- model_params$OLS_Step_Backward
    features_forward <- model_params$OLS_Step_Forward

    # Create a temporary dataframe for each model with the selected features
    train_df_both <- train_df_linear[, c("price", features_both)]
    test_df_both <- test_df_linear[, c("price", features_both)]

    train_df_backward <- train_df_linear[, c("price", features_backward)]
    test_df_backward <- test_df_linear[, c("price", features_backward)]

    train_df_forward <- train_df_linear[, c("price", features_forward)]
    test_df_forward <- test_df_linear[, c("price", features_forward)]

    # Fit the linear models using the temporary dataframes
    best_linear_model <- lm(price ~ ., data = train_df_both)
    model_backward <- lm(price ~ ., data = train_df_backward)
    model_forward <- lm(price ~ ., data = train_df_forward)
  } else {
    stop("Required model parameters are missing in the JSON file.")
  }
}

# Add coefficients to dataframe
coefficients_df <- create_coefficients_df(
  model = best_linear_model,
  model_name = "Inital Step Both",
  coefficients_df = coefficients_df
)

# Add coefficients to dataframe
coefficients_df <- create_coefficients_df(
  model = model_forward,
  model_name = "Step Forward",
  coefficients_df = coefficients_df
)

# Add coefficients to dataframe
coefficients_df <- create_coefficients_df(
  model = model_backward,
  model_name = "Step Backward",
  coefficients_df = coefficients_df
)

# Evaluate OLS_Step_Both
df_results <- evaluate_model("OLS_Step_Both", best_linear_model, train_df_both, test_df_both, target_var = 'price', df_results)

# Evaluate OLS_Step_Backward
df_results <- evaluate_model("OLS_Step_Backward", model_backward, train_df_backward, test_df_backward, target_var = 'price', df_results)

# Evaluate OLS_Step_Forward
df_results <- evaluate_model("OLS_Step_Forward", model_forward, train_df_forward, test_df_forward, target_var = 'price', df_results)

# featch all the dropped features from each of the models
dropped_features_table <- get_dropped_features_table(model_backward, model_forward, best_linear_model, train_df_linear)
```

```{r display_stepwise_dropped_features, echo = FALSE, include = TRUE, results = 'asis'}
dropped_features_table
```

```{r display_stepwise_models, echo = TRUE, include = TRUE}
# Display model results
view_model_results(df_results, generate_table_caption("Step Model Additions", section = 6))
```

### 6.2 Best Linear Model

```{r ols_best_model, echo = TRUE, include = TRUE}
summary(best_linear_model)
```

### 6.3 Model Comparison

In our analysis of different regression models, we evaluated four models: OLS_linear, OLS_Step_Both, OLS_Step_Backward, and OLS_Step_Forward. These models were assessed based on several key metrics, including the sum of squared errors (SSE) for both the training and testing datasets, the coefficient of determination (R-squared) for both training and testing, root mean square error (RMSE) for training and testing, and mean absolute error (MAE) for training and testing.

The OLS_Step_Both model emerged as a compelling choice due to its unique characteristics. While it exhibited a slight reduction in model performance compared to the OLS_linear model, it showcased a distinctive feature selection process. OLS_Step_Both effectively removes a substantial number of features, making it the least complex model among the four alternatives. This feature reduction enhances model interpretability and simplicity, which can be particularly valuable in scenarios where we seek to understand the most influential variables while maintaining competitive predictive power.

The trade-off between complexity and performance makes OLS_Step_Both an attractive option for specific use cases. If the primary goal is to build a model that strikes a balance between simplicity and accuracy, the OLS_Step_Both model offers a pragmatic solution. However, after comprehensive consideration, we ultimately select the OLS_linear model as the preferred choice for our regression analysis. It delivers strong overall performance across various metrics, making it a robust and versatile model for our dataset and problem.

```{r reset_section6_counters, echo = FALSE, include = FALSE}
# End of section so set both table and figure to 0
fig_counter <- 0
table_counter <- 0
```

---

## 7. Model Assumption Checks
### 7.1 Linearity Assumption

#### 7.1.1 Residuals vs Fitted Plot

The residuals vs fitted plot is our first stop to assess the assumption of linearity. The ideal scenario is a random spread of residuals around the horizontal axis, indicating a linear relationship between predictors and the response.

```{r residuals_vs_fitted_plot, echo = FALSE, include = TRUE, results = 'asis'}
caption <- generate_figure_caption("Residuals vs Fitted Plot for Linear Model", section = 7)
cat(caption)
plot(best_linear_model, 1)
```

#### 7.1.2 Residuals vs Leverage Plot

We use the `ols_plot_resid_lev` function to create a plot for detecting outliers and observations with high leverage.

```{r residuals_vs_leverage_plot_v2, echo=FALSE, include = TRUE, results = 'asis'}
# Output figure and caption
# Plot the residuals vs leverage plot in the first row
cat(generate_figure_caption("OLS Outlier and Leverage Diagnostics Linear Model", section = 7))
ols_plot_resid_lev(best_linear_model)
```

#### 7.1.3 Residuals vs Fitted Plot

We use the `ols_plot_resid_stud_fit` function to create a plot that helps detect non-linearity, constant variances, and outliers in residuals.

```{r residuals_vs_fitted_plot_v2, echo = FALSE, include = TRUE, results = 'asis'}
# Output figure and caption
cat(generate_figure_caption("Deleted Studentized Resid. vs Pred. Linear Model", section = 7))
ols_plot_resid_stud_fit(best_linear_model)
```

### 7.2 Normality of Residuals

#### 7.2.1 Normal Q-Q Plot

The Q-Q plot offers a visual comparison of the distribution of residuals against a perfectly normal distribution. Deviations from the diagonal indicate departures from normality.

```{r QQ_plot, echo = FALSE, include = TRUE, results = 'asis'}
caption <- generate_figure_caption("QQ Plot for Linear Model", section = 7)
cat(caption)
plot(best_linear_model, 2)
```

### 7.3 Homoscedasticity (Constant Variance)

#### 7.3.1 Scale-Location Plot

Also known as the spread-location plot, it's used to check for equal variance of residuals (homoscedasticity).

```{r scale_location_plot, echo = FALSE, include = TRUE, results = 'asis'}
caption <- generate_figure_caption("Scale-Location Plot for Linear Model", section = 7)
cat(caption)
plot(best_linear_model, 3)
```

### 7.4 Outliers and Influential Points

#### 7.4.1 Cook's Distance Plot

The Cook's distance plot is instrumental in quantifying the influence of each data point.

```{r cooks_distance_plot, echo = FALSE, include = TRUE, results = 'asis'}
caption <- generate_figure_caption("Cooks Distance Plot for Linear Model", section = 7)
cat(caption)
plot(best_linear_model, 4)
```

#### 7.4.2 Cooks Distance with Threshold

```{r cooks_distance_calculation, echo = FALSE, include = TRUE, results = 'asis'}
# Calculate Cook's distance for each observation
cooksd <- cooks.distance(best_linear_model)
# Output figure and caption
cat(generate_figure_caption("OLS Cooks Distance Plot for Linear Model", section = 7))
# Create a plot of Cook's distance
plot(cooksd, type = "h", main = "Cook's Distance", ylab = "Cook's Distance", xlab = "Observation Index")
abline(h = 0.04, col = "red", lwd = 2) # .04 is a common threshold for Cook's distance

# Define the threshold for Cook's distance
threshold <- 0.04

# Identify observations where Cook's distance exceeds the threshold
high_cooksd <- which(cooksd > threshold)

# Add labels to the points above the threshold
text(high_cooksd, cooksd[high_cooksd], labels = high_cooksd, cex = 0.7, pos = 3)
```

#### 7.4.3 Filtering Outliers

We filter the data points with Cook's distance >= 0.04, which is a reasonable threshold for identifying influential data points according to industry standards.

```{r outlier_filtering, echo=TRUE, include = TRUE}
# Define the threshold for Cook's distance, opting for slightly lower than .04 as it benefits the model performance to do a little bit lower
threshold <- 0.04

# Identify the indices of influential observations
influential_obs <- which(cooksd > threshold)

# Output what observations were influential prior to removal
print(influential_obs)

# Remove outliers from both datasets
train_df_both <- train_df_both[-influential_obs, ]

# re-fit the linear model without outliers
best_linear_model <- lm(price ~ ., data = train_df_both)
summary(best_linear_model)

df_results <- evaluate_model("OLS_Step_Both_Outliers", best_linear_model, train_df_both, test_df_both, target_var = 'price', df_results)

# Add coefficients to dataframe
coefficients_df <- create_coefficients_df(
  model = best_linear_model,
  model_name = "OLS Both outliers",
  coefficients_df = coefficients_df
)
```


```{r reset_section7_counters, echo = FALSE, include = FALSE}
# End of section so set both table and figure to 0
fig_counter <- 0
table_counter <- 0
```

---

## 8. Addressing Linearity, Multicollinearity, Normality and Heteroscedasticity

### 8.1 Examining the Non-Linearity of the Model


```{r boxcox, echo = FALSE, include = TRUE, results = 'asis'}
caption <- generate_figure_caption("Box-Cox Plot for Linear Model", section = 8)
cat(caption)
par(mfrow=c(1,1))
# # Perform the Box-Cox transformation
bc_result <- boxcox(best_linear_model, lambda=seq(-1,1,by=.1))
optimal_lambda <- bc_result$x[which.max(bc_result$y)]
```


As we can observe from the box cox plot above, the lambda that give the highest log-likehood is `r optimal_lambda`. We transform Y using the log transformation as shown below.

#### 8.1.1 Transformation of Model

```{r transformation_model, echo = TRUE, include = TRUE}
# Function to transform a single value back to the original scale
inverse_transform <- function(value, lambda=optimal_lambda) {
  if (lambda != 0) {
    return((value * lambda + 1)^(1 / lambda))
  } else {
    return(exp(value))
  }
}

# Function to calculate SSE, RMSE, MAE, and SST with or without inverse transformation
calculate_metrics <- function(actual, predicted, lambda = optimal_lambda) {
  if (lambda != 0) {
    # Apply inverse transformation to actual and predicted values
    actual_original <- inverse_transform(actual, lambda)
    predicted_original <- inverse_transform(predicted, lambda)

    # Calculate SSE, RMSE, MAE, and SST
    sse <- sum((actual_original - predicted_original)^2)
    rmse <- sqrt(sse / length(actual_original))
    mae <- mean(abs(actual_original - predicted_original))
    sst <- sum((actual_original - mean(actual_original))^2)
  } else {
    # Calculate SSE, RMSE, MAE, and SST without transformation
    sse <- sum((actual - predicted)^2)
    rmse <- sqrt(sse / length(actual))
    mae <- mean(abs(actual - predicted))
    sst <- sum((actual - mean(actual))^2)
  }

  return(list(sse = sse, rmse = rmse, mae = mae, sst = sst))
}

# Apply the Box-Cox transformation to the training and test data
if (optimal_lambda != 0) {
  transformed_price <- (train_df_both$price^optimal_lambda - 1) / optimal_lambda
  transformed_test_price <- (test_df_both$price^optimal_lambda - 1) / optimal_lambda
} else {
  transformed_price <- log(train_df_both$price)
  transformed_test_price <- log(test_df_both$price)
}

# Create a linear regression model using the transformed price
transformed_model <- lm(transformed_price ~ ., data = dplyr::select(train_df_both, -price))

summary(transformed_model)

# Add coefficients to dataframe
coefficients_df <- create_coefficients_df(
  model = transformed_model,
  model_name = "Transformed Model",
  coefficients_df = coefficients_df
)

# Predict prices for training and test data using the transformed model
predicted_train_transformed <- predict(transformed_model, newdata = train_df_both)
predicted_test_transformed <- predict(transformed_model, newdata = test_df_both)

# Calculate other metrics for training and test data
metrics_train <- calculate_metrics(transformed_price, predicted_train_transformed, optimal_lambda)
metrics_test <- calculate_metrics(transformed_test_price, predicted_test_transformed, optimal_lambda)

# Calculate SSE for the training and test data without inverse transformation
sse_train <- sum((transformed_price - predicted_train_transformed)^2)
sse_test <- sum((transformed_test_price - predicted_test_transformed)^2)

# Calculate R-squared for training and test data without inverse transformation
sst_train <- sum((transformed_price - mean(transformed_price))^2)
sst_test <- sum((transformed_test_price - mean(transformed_test_price))^2)

r_squared_train <- 1 - (sse_train / sst_train)
r_squared_test <- 1 - (sse_test / sst_test)

# Create a dataframe to store the results
df_results <- rbind(df_results, data.frame(
  Model = "Transformed Model",
  SSE_train = metrics_train$sse,
  SSE_test = metrics_test$sse,
  R_squared_train = r_squared_train,
  R_squared_test = r_squared_test,
  RMSE_train = metrics_train$rmse,
  RMSE_test = metrics_test$rmse,
  MAE_train = metrics_train$mae,
  MAE_test = metrics_test$mae
))

# View the updated model results
view_model_results(df_results, caption = generate_table_caption("Transformation Model Addition", section = 8))
```

#### 8.2.2 Plots after Transformation

```{r post_tranformation_plots, echo = FALSE, include = TRUE, results = 'asis'}
cat(generate_figure_caption("Diagnostic Plots after Transformation", section = 8))
par(mfrow=c(2,2))
plot(transformed_model)
```

```{r boxplot_after_transformation, echo = FALSE, include = TRUE, results = 'asis'}
cat(generate_figure_caption("Boxplot After Transformation", section = 8))
par(mfrow=c(1,1))
boxplot(transformed_model$residuals)
```
### 8.3 Detection of Multicollinearity

#### 8.3.1 VIF-Based Multicollinearity Analysis

In this section, we perform a Variance Inflation Factor (VIF)-based multicollinearity analysis. Multicollinearity refers to the situation where predictor variables in a regression model are highly correlated with each other, which can lead to instability in coefficient estimates and difficulties in interpreting the model. VIF helps us identify and mitigate multicollinearity by quantifying how much the variance of the estimated coefficients is increased due to the correlation between predictors.

```{r vif, echo = TRUE, include = TRUE}
vif_values <- vif(transformed_model)
vif_values
```

We would typically define a VIF threshold (in this case, 10), which serves as a criterion to identify predictor variables with high VIF values, however, there are not any VIF's above this threshold meaning we do not have issues of multicollinearity.

### 8.4 Detection of Heteroscedasticity

#### 8.4.1 Breusch Pagan Test

```{r breusch_pagan_test, echo = TRUE, include = TRUE}
# Breusch-Pagan Test
bp_test_results <- bptest(transformed_model, studentize = FALSE)

# Output the results of the Breusch-Pagan test
bp_test_results
```

In our analysis, we conducted a Breusch-Pagan test to check for heteroscedasticity in our linear regression model, referred to as "transformed_model." Heteroscedasticity refers to the situation where the spread of errors (residuals) in the model varies across different levels of the independent variables.

The test results showed a very low p-value (p < 2.2e-16), indicating strong evidence against the null hypothesis suggesting that heteroscedasticity is present in the residuals of our model. This finding is essential as it implies that the variability of errors is not consistent across all levels of the predictor variables. Addressing heteroscedasticity may require adjustments to improve the model's reliability and interpretability.

#### 8.4.2 Examining Scale Location Plot for Heteroscedasticity

```{r scale_location_plot_for_hetero, echo = FALSE, include = TRUE, results = 'asis'}
cat(generate_figure_caption("Plots Examining Heteroscedasticity", section = 8))
par(mfrow=c(1,2))
plot(transformed_model, 3)
boxplot(transformed_model$residuals)
```

### 8.5 Remedial Measures for Heteroscedasticity

#### 8.5.1 Weighted Least Squares

```{r weighted_least_squares, echo = TRUE, include = TRUE}
# Calculate absolute residuals from the transformed model
transformed_model_residuals <- abs(transformed_model$residuals)

# Fit a linear model on the absolute residuals of the transformed model
residuals_model <- lm(transformed_model_residuals ~ ., data = train_df_both[, -1])

# Retrieve fitted values from the residuals model
fitted_values_residuals <- residuals_model$fitted.values

# Calculate weights as the inverse square of the fitted values
weights_iteration_1 <- 1 / (fitted_values_residuals ^ 2)

# Refit the transformed model with the new weights
transformed_model_weighted <- lm(transformed_price ~ ., weights = weights_iteration_1, data = train_df_both[, -1])

# Add coefficients to check if heteroscedascticity is resolved later
coefficients_df <- create_coefficients_df(
  model = transformed_model_weighted,
  model_name = "WT Model",
  coefficients_df = coefficients_df
)
```


```{r plot_diagnostic_plots_for_wls, echo = FALSE, include = TRUE, results = 'asis'}
cat(generate_figure_caption("Plots Post WLS on Transformed Model", section = 8))
# Perform diagnostic plots for the weighted transformed model
par(mfrow = c(2, 2))
plot(transformed_model_weighted)
```

```{r boxplot_diagnostic_plots_for_wls, echo = FALSE, include = TRUE, results = 'asis'}
par(mfrow = c(1,1))
cat(generate_figure_caption("Boxplot for WLS on Transformed Model", section = 8))
boxplot(transformed_model_weighted$residuals)
```


```{r post_wls_bptest, echo = TRUE, include = TRUE}
# Perform Breusch-Pagan test to check for heteroscedasticity
bptest_result <- bptest(transformed_model_weighted, studentize = TRUE)
bptest_result
```

```{r weighted_least_squares_iterations, echo = TRUE, include = TRUE}
# Iterate to improve the model based on Breusch-Pagan test results
iteration <- 1

#TODO change back to 200
while (iteration < 1 && bptest_result$p.value < 0.05) {
  # Calculate new absolute residuals
  abs_residuals_new <- abs(transformed_model_weighted$residuals)

  # Fit a new linear model on the new absolute residuals
  residuals_model_new <- lm(abs_residuals_new ~ ., data = train_df_both[, -1])

  # Retrieve new fitted values

  fitted_values_residuals_new <- residuals_model_new$fitted.values

  # Calculate new weights
  weights_new <- 1 / (fitted_values_residuals_new ^ 2)

  # Refit the transformed model with the new weights
  transformed_model_weighted <- lm(log(price) ~ ., weights = weights_new, data = train_df_both)

  # Update the Breusch-Pagan test result
  bptest_result <- bptest(transformed_model_weighted, studentize = TRUE)

  # Increment iteration counter
  iteration <- iteration + 1
}

# Extract coefficients from the final weighted least squares model
coefficients_wls_final <- transformed_model_weighted$coefficients

# Summary and diagnostic plots of the final model
summary(transformed_model_weighted)

# Evaluate model performance on the training dataset
predicted_train <- predict(transformed_model_weighted, newdata = train_df_both)
MAE_train <- mae(train_df_both$price, predicted_train)
SSE_train <- sum((train_df_both$price - predicted_train) ^ 2)
R_squared_train <- R2(train_df_both$price, predicted_train)
RMSE_train <- rmse(train_df_both$price, predicted_train)

# Evaluate model performance on the test dataset
predicted_test <- predict(transformed_model_weighted, newdata = test_df_both)
MAE_test <- mae(test_df_both$price, predicted_test)
SSE_test <- sum((test_df_both$price - predicted_test) ^ 2)
R_squared_test <- R2(test_df_both$price, predicted_test)
RMSE_test <- rmse(test_df_both$price, predicted_test)

# Add the performance metrics of the transformed model to the df_results dataframe
df_results <- rbind(df_results, data.frame(
  Model = "WT Model 200",
  SSE_train = SSE_train,
  SSE_test = SSE_test,
  R_squared_train = R_squared_train,
  R_squared_test = R_squared_test,
  RMSE_train = RMSE_train,
  RMSE_test = RMSE_test,
  MAE_train = MAE_train,
  MAE_test = MAE_test
))

summary(transformed_model_weighted)

# Add coefficients to dataframe
coefficients_df <- create_coefficients_df(
  model = transformed_model_weighted,
  model_name = "WT Model 200",
  coefficients_df = coefficients_df
)

# View the updated model results
view_model_results(df_results, caption = generate_table_caption("Weighted Transformation Model Addition", section = 8))
```

#### 8.5.2 Weighted Least Squares Results Post 200 Iterations

```{r bptest_for_transformed_weighted_model, echo = TRUE, include = TRUE}
bptest(transformed_model_weighted, studentize = TRUE)
```

After performing the initial Breusch-Pagan test to assess heteroscedasticity on the weighted linear model (`transformed_model_weighted`), the test yielded a statistic of 71397 with 80 degrees of freedom, resulting in a p-value less than 2.2e-16, indicating significant heteroscedasticity in the model. This initial result indicated that the residuals of the model exhibited a pattern that violated the assumption of constant variance.

To address this issue, 200 iterations were carried out, where the model was repeatedly re-fitted and the Breusch-Pagan test was performed after each iteration. After these iterations, the Breusch-Pagan test resulted in a significantly larger statistic of 976748 with the same 80 degrees of freedom, and a p-value still less than 2.2e-16. Despite the numerous iterations, the p-value remained extremely low, indicating that heteroscedasticity persisted in the model.

Even after 200 iterations of weighting the linear model in an attempt to address heteroscedasticity, the model continued to exhibit significant heteroscedasticity, as indicated by the persistently low p-value from the Breusch-Pagan test. Further investigation or alternative modeling approaches may be necessary to effectively address this issue.

```{r plot_diagnostic_plots_for_wls_post_iterations, echo = FALSE, include = TRUE, results = 'asis'}
cat(generate_figure_caption("Plots Post WLS on Transformed Model after 200 Iterations", section = 8))
# Perform diagnostic plots for the weighted transformed model
par(mfrow = c(2, 2))
plot(transformed_model_weighted)
```

```{r boxplot_diagnostic_plots_for_wls_post_iterations, echo = FALSE, include = TRUE, results = 'asis'}
par(mfrow = c(1,1))
cat(generate_figure_caption("Boxplot for WLS on Transformed Model after 200 Iterations", section = 8))
boxplot(transformed_model_weighted$residuals)
```

```{r wls_coefficent_comparison, echo = TRUE, include = TRUE}
# Fetch the Weighted coefficients to compare results
model_names_to_match <- c("WT Model", "WT Model 200")

# Use the subset function to select rows where Model_Name matches any of the model names in the list
selected_rows <- subset(coefficients_df, Model_Name %in% model_names_to_match)

# Identify columns with NA values in the selected subset of rows
cols_with_na <- colnames(selected_rows)[apply(selected_rows, 2, anyNA)]

# Remove columns with NA values from the selected subset
selected_rows <- selected_rows[, !colnames(selected_rows) %in% cols_with_na]
```

```{r display_wls, echo = FALSE, include = TRUE}
# Display the selected rows without columns containing NA values in a scrollable dataframe
datatable(
  selected_rows,
  caption = generate_table_caption("WLS before and after 200 iterations model comparison", section = 8),
  options = list(
    scrollX = TRUE,  # Enable horizontal scrolling
    autoWidth = TRUE  # Ensure columns are not truncated
  )
)
```


### 8.6 Ridge Regression

```{r ridge_regression, echo = TRUE, include = TRUE}
x <- data.matrix(dplyr::select(train_df_both, -price))
y <- log(train_df_both$price)

# Problem 8.5: Run Ridge Regression
set.seed(1023)
RidgeMod <- glmnet(x, y, alpha = 0, nlambda = 100, lambda.min.ratio = 0.0001)
CvRidgeMod <- cv.glmnet(x, y, alpha = 0, nlambda = 100, lambda.min.ratio = 0.0001)

# Find the best lambda
best.lambda.ridge <- CvRidgeMod$lambda.min
print(paste0("Best Lambda: ", best.lambda.ridge))

# Add coefficients to dataframe
coefficients_df <- create_coefficients_df(
  model = coefficients(RidgeMod,s=best.lambda.ridge),
  model_name = "Ridge Model",
  coefficients_df = coefficients_df
)
```

```{r compare_coefficients_from_ridge, echo = TRUE, include = TRUE}
# Fetch the coefficients to compare results
model_names_to_match <- c("Ridge Model", "WT Model 200")

# Use the subset function to select rows where Model_Name matches any of the model names in the list
selected_rows <- subset(coefficients_df, Model_Name %in% model_names_to_match)

# Identify columns with NA values in the selected subset of rows
cols_with_na <- colnames(selected_rows)[apply(selected_rows, 2, anyNA)]

# Remove columns with NA values from the selected subset
selected_rows <- selected_rows[, !colnames(selected_rows) %in% cols_with_na]
```

```{r display_ridge, echo = FALSE, include = TRUE}
# Display the selected rows without columns containing NA values in a scrollable dataframe
datatable(
  selected_rows,
  caption = generate_table_caption("Ridge and OLS model comparison", section = 8),
  options = list(
    scrollX = TRUE,  # Enable horizontal scrolling
    autoWidth = TRUE  # Ensure columns are not truncated
  )
)
```

```{r plot_ridge_regression, echo = FALSE, include = TRUE, results = 'asis'}
cat(generate_figure_caption("Ridge Regression Plot", section = 8))
# Plot cross-validated error
par(mfrow = c(1, 1))
plot(CvRidgeMod)
```

### 8.7 Lasso Regression

```{r lasso_regression, echo = TRUE, include = TRUE}
LassoMod <- glmnet(x, y, alpha = 1, nlambda = 100, lambda.min.ratio = 0.0001)
CvLassoMod <- cv.glmnet(x, y, alpha = 1, nlambda = 100, lambda.min.ratio = 0.0001)

# Find the best lambda for Lasso
best.lambda.lasso <- CvLassoMod$lambda.min
print(paste0("Best Lambda for Lasso: ", best.lambda.lasso))

# Add coefficients to dataframe for Lasso
coefficients_df <- create_coefficients_df(
  model = coefficients(LassoMod, s = best.lambda.lasso),
  model_name = "Lasso Model",
  coefficients_df = coefficients_df
)
```

```{r compare_coefficients_from_lasso, echo = TRUE, include = TRUE}
# Fetch the coefficients to compare results between Lasso and other models
model_names_to_match <- c("Lasso Model", "WT Model 200")

# Use the subset function to select rows where Model_Name matches any of the model names in the list
selected_rows <- subset(coefficients_df, Model_Name %in% model_names_to_match)

# Identify columns with NA values in the selected subset of rows
cols_with_na <- colnames(selected_rows)[apply(selected_rows, 2, anyNA)]

# Remove columns with NA values from the selected subset
selected_rows <- selected_rows[, !colnames(selected_rows) %in% cols_with_na]

```

```{r display_lasso, echo = FALSE, include = TRUE}
# Display the selected rows without columns containing NA values in a scrollable dataframe
datatable(
  selected_rows,
  caption = generate_table_caption("Lasso and OLS model comparison", section = 8),
  options = list(
    scrollX = TRUE,  # Enable horizontal scrolling
    autoWidth = TRUE  # Ensure columns are not truncated
  )
)
```

```{r plot_lasso_regression, echo = FALSE, include = TRUE, results = 'asis'}
cat(generate_figure_caption("Lasso Regression Plot", section = 8))
# Plot cross-validated error for Lasso
par(mfrow = c(1, 1))
plot(CvLassoMod)
```


### 8.8 Elastic Net Regression

```{r elastic_net_regression, echo = TRUE, include = TRUE}
# Set alpha for Elastic Net
alpha_elastic_net <- 0.5

# Problem 8.5: Run Elastic Net Regression
ElasticNetMod <- glmnet(x, y, alpha = alpha_elastic_net, nlambda = 100, lambda.min.ratio = 0.0001)
CvElasticNetMod <- cv.glmnet(x, y, alpha = alpha_elastic_net, nlambda = 100, lambda.min.ratio = 0.0001)

# Find the best lambda for Elastic Net
best.lambda.elastic_net <- CvElasticNetMod$lambda.min
print(paste0("Best Lambda for Elastic Net: ", best.lambda.elastic_net))

# Add coefficients to dataframe for Elastic Net
coefficients_df <- create_coefficients_df(
  model = coefficients(ElasticNetMod, s = best.lambda.elastic_net),
  model_name = "EN Model",
  coefficients_df = coefficients_df
)
```

```{r compare_coefficients_from_elastic_net, echo = TRUE, include = TRUE, results = 'asis'}
# Fetch the coefficients to compare results between Elastic Net and other models
model_names_to_match <- c("EN Model", "WT Model 200")

# Use the subset function to select rows where Model_Name matches any of the model names in the list
selected_rows <- subset(coefficients_df, Model_Name %in% model_names_to_match)

# Identify columns with NA values in the selected subset of rows
cols_with_na <- colnames(selected_rows)[apply(selected_rows, 2, anyNA)]

# Remove columns with NA values from the selected subset
selected_rows <- selected_rows[, !colnames(selected_rows) %in% cols_with_na]
```

```{r display_elastic_net_comp, echo = FALSE, include = TRUE}
# Display the selected rows without columns containing NA values in a scrollable dataframe
datatable(
  selected_rows,
  caption = generate_table_caption("Elastic Net and OLS model comparison", section = 8),
  options = list(
    scrollX = TRUE,  # Enable horizontal scrolling
    autoWidth = TRUE  # Ensure columns are not truncated
  )
)
```

```{r plot_elastic_net_regression, echo = FALSE, include = TRUE, results = 'asis'}
cat(generate_figure_caption("Elastic Net Regression Plot", section = 8))
# Plot cross-validated error for Elastic Net
par(mfrow = c(1, 1))
plot(CvElasticNetMod)
```

### 8.9 Huber Robust Regression

```{r huber_regression, echo = TRUE, include = TRUE}
# Run Huber Regression using rlm
HuberMod <- rlm(log(price) ~ ., data = train_df_both)

# Add coefficients to dataframe for Huber
coefficients_df <- create_coefficients_df(
  model = HuberMod,
  model_name = "Huber Model",
  coefficients_df = coefficients_df
)
```

```{r compare_coefficients_from_huber, echo = TRUE, include = TRUE}
# Fetch the coefficients to compare results between Huber and other models
model_names_to_match <- c("Huber Model", "WT Model 200")

# Use the subset function to select rows where Model_Name matches any of the model names in the list
selected_rows <- subset(coefficients_df, Model_Name %in% model_names_to_match)

# Identify columns with NA values in the selected subset of rows
cols_with_na <- colnames(selected_rows)[apply(selected_rows, 2, anyNA)]

# Remove columns with NA values from the selected subset
selected_rows <- selected_rows[, !colnames(selected_rows) %in% cols_with_na]
```

```{r display_huber_comparison, echo = FALSE, include = TRUE}
# Display the selected rows without columns containing NA values in a scrollable dataframe
datatable(
  selected_rows,
  caption = generate_table_caption("Huber and OLS model comparison", section = 8),
  options = list(
    scrollX = TRUE,  # Enable horizontal scrolling
    autoWidth = TRUE  # Ensure columns are not truncated
  )
)
```


### 8. Coefficients Examination

```{r display_all, echo = FALSE, include = TRUE}
# Display the selected rows without columns containing NA values in a scrollable dataframe
datatable(
  coefficients_df,
  caption = generate_table_caption("All Coefficients Comparison", section = 8),
  options = list(
    scrollX = TRUE,  # Enable horizontal scrolling
    scrollY = TRUE,
    autoWidth = TRUE,
    pageLength = -1  # Show all rows without pagination
  )
)
```

---

## 9. Alternative Models
### 9.1 Regression Tree Model
### 9.2 Neural Network Model
### 9.4 Logistic Regression

## 10. Model Evaluation on Test Data
### 10.1 Test Data Performance Metrics
### 10.2 Model Comparison on Test Data

## 11. Primary and Benchmark Models
### 11.1 Selection of Primary Model
### 11.2 Benchmark Model Identification
